# Results

Field assessments were conducted between `r format(min(pscis$date), format="%B %d %Y")` and `r format(max(pscis$date), format="%B %d %Y")` by Allan Irvine, R.P.Bio, Kyle Prince, P,Biol and Mark Fjeld, BiT.  

## Phase 1

A total of `r n_distinct(pscis$my_crossing_reference)` phase 1 assessments were conducted between with site details and photos presented in [Appendix - Phase 1 Fish Passage Assessment Data and Photos].  A cost benefit analysis for phase 1 sites is summarized in Table \@ref(tab:cost-est-phase-1).

<br>

```{r cost-est-phase-1}
tab_cost_est_phase1 %>% 
  my_kable_scroll(caption_text = 'Modelled upstream habitat estimate and cost benefit for Phase 1 assessments.')
  # kable(caption = 'Modelled upstream habitat estimate and cost benefit.',
  #       escape = T) %>% 
  # kableExtra::kable_styling(c("condensed"), full_width = T, font_size = 11) %>% 
  # kableExtra::scroll_box(width = "100%", height = "500px")

```


## Phase 2

 
Habitat confirmation assessments were conducted at `r hab_site_priorities %>% nrow()` with a total of approximately 15 km of stream assessed. `r hab_site_priorities %>% filter(priority %ilike% 'high') %>% nrow() %>% english::as.english() %>% str_to_title()` crossings were rated as high priorities for remediation design, `r hab_site_priorities %>% filter(priority %ilike% 'Moderate') %>% nrow() %>% english::as.english() %>% str_to_title()` crossings rated as moderate priorities, `r hab_site_priorities %>% filter(priority %ilike% 'Low') %>% nrow() %>% english::as.english()` crossing was rated as a low priority and `r hab_site_priorities %>% filter(priority %ilike% 'no fix') %>% nrow() %>% english::as.english()` rated as "no fix". 

<br>


Raw results are included in digital format as [Attachment 2](https://github.com/NewGraphEnvironment/fish_passage_elk_2020_reporting/raw/master/data/habitat_confirmations.xls) and summarized in Tables \@ref(tab:tab-overview) - \@ref(tab:tab-habitat-summary)


<br>

```{r tab-overview}
# tab_overview %>% 
#   kable(caption = 'Overview of habitat confirmation sites.') %>% 
#   kableExtra::kable_styling(c("condensed"), full_width = T, font_size = 11) %>% 
#   kableExtra::scroll_box(width = "100%", height = "500px")

tab_overview %>% 
  my_kable_scroll(caption_text = 'Overview of habitat confirmation sites.')
```

<br>

```{r}
make_tab_cv(dat = pscis2) %>% 
  my_kable(caption_text = 'Summary of Phase 2 fish passage reassessments.')
```

<br>

```{r cost-est-phase-2}
tab_cost_est_phase2 %>% 
  my_kable_scroll(caption_text = 'Cost benefit analysis for Phase 2 assessments.')
  # kable(caption = 'Modelled upstream habitat estimate and cost benefit.',
  #       escape = T) %>% 
  # kableExtra::kable_styling(c("condensed"), full_width = T, font_size = 11) %>% 
  # kableExtra::scroll_box(width = "100%", height = "500px")

```

<br>

```{r tab-habitat-summary}
tab_hab_summary %>% 
  filter(Location %ilike% 'upstream') %>% 
  select(-Location) %>% 
  my_kable(caption_text = 'Summary of Phase 2 habitat confirmation details.')

```

