---
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup-50261, eval = F}
knitr::opts_chunk$set(echo=FALSE, message=FALSE, warning=FALSE)
source('R/packages.R')
source('R/functions.R')
source('R/functions-phase2.R')
source('R/tables-phase2.R')
source('R/tables.R')
source('R/extract-fish.R')
source('R/functions-fish.R')


```



```{r  load-data-50261}
my_site <- 50261

```


```{r eval = F}
my_bcfish <- my_bcfishpass() ##can be useful to look at

# `r paste0('# Appendix - Crossing ', my_site)`{-} 
# `r paste0('**', my_overview_info() %>% pull(road_name), ' - ', my_overview_info() %>% pull(stream_name), '**')`
```

`r appendix_title()`
`r appendix_subtitle()`



## Site Location {-}

Crossing `r as.character(my_site)` is located on a `r my_pscis_info() %>% pull(stream_name)`, approximately `r my_priority_info(loc = 'ds') %>% pull(survey_length_m)`m upstream from the confluence with the Michel Creek. The crossing is located at 7.6km on the `r my_overview_info() %>% pull(road_name)` accessed from Coal Mountain. 


<br>

## Background {-}

At the crossing location, the stream is `r my_bcfishpass() %>% pull(stream_order) %>% english::ordinal()` order with a watershed area upstream of the road of approximately `r my_bcfishpass() %>% mutate(watershed_upstr_ha = watershed_upstr_ha/100) %>% pull(watershed_upstr_ha) %>% round(1)` km^2^. The elevation of the watershed ranges from a maximum of 2600 to 1700m at the culvert. There is one crossing modelled downstream of PCSIS `r as.character(my_site)` (4600837) and one modelled above (4602097). There was no fisheries information available for the watershed [@moeStreamInventorySample;@data_fish_obs]. 

<br>

PSCIS stream crossing `r as.character(my_site)` was ranked as a moderate priority for follow up with habitat confirmation due to the relatively large amount of potential habitat modelled as suitable for westslope cutthrout trout occupancy (<20% for ~3km) located upstream and a medium habitat value ranking by @grainger2011FishPassage2011. The habitat confirmation was completed on `r my_pscis_info() %>% pull(date) %>% format(., format = "%B %d, %Y")`. A map of the watershed including areas surveyed is provided in Attachment 1 – Map [`r my_bcfishpass() %>% pull(map_tile_display_name)`](`r my_mapsheet()`).

<br>

## Stream Characteristics at Crossing {-}

At the time of the survey, the three un-embedded and non-backwatered `r my_pscis_info() %>% pull(diameter_or_span_meters)`m diameter crossings were considered a barrier to upstream fish passage with a pipe lengths of `r my_pscis_info() %>% pull(length_or_width_meters)`m, culvert slopes of `r my_pscis_info() %>% pull(culvert_slope_percent)`%, a stream width ratio of `r my_pscis_info() %>% pull(stream_width_ratio)` and outlet drops of `r my_pscis_info() %>% pull(outlet_drop_meters)`m (Table \@ref(tab:tab-culvert-50261)). Water temperature was `r my_habitat_info() %>% filter(rowname == 'temperature c') %>% pull(us)`$^\circ$C, pH was `r my_habitat_info() %>% filter(rowname == 'p h') %>% pull(us)` and conductivity was `r my_habitat_info() %>% filter(rowname == 'conductivity m s cm') %>% pull(us)`uS/cm.  

<br>


```{r eval=F}
##this is useful to get some comments for the report
hab_site %>% filter(site == my_site & location == 'ds') %>% pull(comments)
my_hab_info <- my_habitat_info()
```

## Stream Characteristics Downstream {-}

The stream was surveyed downstream from the culvert for `r my_priority_info(loc = 'ds') %>% pull(survey_length_m)`m. Overall total cover amount was rated as `r my_habitat_info2(loc = 'ds') %>% filter(rowname == 'total cover') %>% pull(v)` with  `r my_habitat_info2(loc = 'ds') %>% filter(v == 'dominant') %>% pull(rowname)` dominant. Cover was also present as `r my_habitat_info2(loc = 'ds') %>% filter(v == 'sub-dominant') %>% pull(rowname) %>% knitr::combine_words()` (Table \@ref(tab:tab-habitat-summary-50261), Figure \@ref(fig:photo-50261-01)). The average channel width was `r my_habitat_info3(loc = 'ds', row = 'avg channel width m')`m, the average wetted width was `r my_habitat_info3(loc = 'ds', row = 'avg wetted width m')`m and the average gradient was `r my_habitat_info3(loc = 'ds', row = 'average gradient percent')`%.  The dominant substrate was `r my_habitat_info3(loc = 'ds', row = "bed material dominant")` with `r my_habitat_info3(loc = 'ds', row = "bed material subdominant")` subdominant.  Some intermittedt pools were present with occasional pockets of gravel suitable for spawning.  A 4.4m high chute was located 200m downstream of crossing (UTM: 11U 668858 5481210) and is considered a permanent impassable barrier to upstream migration (Figure \@ref(fig:photo-50261-01b).  

<br>



```{r eval=F}
##this is useful to get some comments for the report
hab_site %>% filter(site == my_site & location == 'us') %>% pull(comments)
my_hab_info <- my_habitat_info()
```


<br>

## Stream Characteristics Upstream {-}

The stream was surveyed immediately upstream from the culvert for `r my_priority_info(loc = 'ds') %>% pull(survey_length_m)`m. Overall, total cover amount was rated as `r my_habitat_info2(loc = 'us') %>% filter(rowname == 'total cover') %>% pull(v)` with  `r my_habitat_info2(loc = 'us') %>% filter(v == 'dominant') %>% pull(rowname)` dominant. Cover was also present as `r my_habitat_info2(loc = 'us') %>% filter(v == 'sub-dominant') %>% pull(rowname) %>% knitr::combine_words()` (Table \@ref(tab:tab-habitat-summary-50261), Figure \@ref(fig:photo-50261-02)). The average channel width was `r my_habitat_info3(loc = 'us', row = 'avg channel width m')`m, the average wetted width was `r my_habitat_info3(loc = 'us', row = 'avg wetted width m')`m and the average gradient was `r my_habitat_info3(loc = 'us', row = 'average gradient percent')`%.  The dominant substrate was `r my_habitat_info3(loc = 'us', row = "bed material dominant")` with `r my_habitat_info3(loc = 'us', row = "bed material subdominant")` subdominant.  Overall, habitat was rated as `r my_priority_info(loc = 'ds') %>%  pull(hab_value)` value.  There were multiple natural barriers to upstream fish passage within the area surveyed that were considered too steep for upstream migration by any species.

<br>


## Conclusion {-}

The chute located 200m downstream of PSCIS `r as.character(my_site)` is an impassable barrier to upstream migration and negates the value in restoring passage at the crossing. It is recommended that the crossing not proceed to design for remeiation of fish passage at the site. 

\pagebreak

```{r tab-culvert-50261, eval = T}
print_tab_summary()

```

\pagebreak

```{r tab-habitat-summary-50261}
tab_hab_summary %>% 
  filter(Site  == my_site) %>% 
  select(-Site) %>%
  my_kable(caption_text = paste0('Summary of habitat details for PSCIS crossing ', my_site, '.'))

```



<br>

```{r photo-50261-01, fig.cap= my_caption, out.width = photo_width, eval=T}
my_photo = pull_photo_by_str(str_to_pull = '_d1')

my_caption = paste0('Habitat downstream of PSCIS crossing ', my_site, '.')

grid::grid.raster(get_img())

# knitr::include_graphics(get_img_path())
```

<br>

```{r photo-50261-01b, fig.cap= my_caption, out.width = photo_width, eval=T}
my_photo = pull_photo_by_str(str_to_pull = 'chute')

my_caption = paste0('Chute barrier located 200 downstream of PSCIS crossing ', my_site, '.')

grid::grid.raster(get_img())

# knitr::include_graphics(get_img_path())
```

<br>

```{r photo-50261-02, fig.cap= my_caption, out.width = photo_width, eval=T}
my_photo = pull_photo_by_str(str_to_pull = '_u1')

my_caption = paste0('Habitat upstream of PSCIS crossing ', my_site, '.')

grid::grid.raster(get_img())

# knitr::include_graphics(get_img_path())
```


