
```{r setup-50152, eval = F}
knitr::opts_chunk$set(echo=FALSE, message=FALSE, warning=FALSE)
source('R/packages.R')
source('R/functions.R')
source('R/functions-phase2.R')
source('R/tables-phase2.R')
source('R/tables.R')
source('R/extract-fish.R')
source('R/functions-fish.R')


```



```{r  load-data-50152}
my_site <- 50152
```


`r paste0('# Appendix - Crossing ', my_site)`{-} 
`r paste0('**', my_overview_info() %>% pull(road_name), ' - ', my_overview_info() %>% pull(stream_name), '**')`


## Site Location {-}

Crossing `r as.character(my_site)` is located on a `r my_pscis_info() %>% pull(stream_name)`, approximately `r my_priority_info(loc = 'ds') %>% pull(survey_length_m)`m upstream from the confluence with Lizard Creek and within Mt.Fernie provincial Park. This culvert has also been recorded in PSCIS as crossing 62521. The culvert is located on a paved section of Mt.Fernie Park Road (0.3km mark) and is accessed from Highway 3 within Fernie city limits.  The Mt.Fernie Provincial Park is a popular recreational destination for hikers, sightseers and mountain bikers.  Island Lake Lodge is located at 1400m of elevation near Island Lake and is a year round tourist destination providing accommodations, guided hiking and backcountry catskiing in the Lizard Creek watershed. 


<br>

## Background {-}

The subject culvert appears to have been recently replaced with a new structure.  No fisheries information was available for the stream although Lizard Creek supports westslope cutthroat trout, bull trout, mountain whitefish, brook trout, longnose sucker and longnose dace downstream [@moeStreamInventorySample; @data_fish_obs]. @elkriveralliance2020ElkRiver conducted redd surveys in Lizard Creek in 2019 along with Morrissey Creek, Coal Creek and Forsyth Creek.  A total of 55 redds were observed within a 2.4km section of Lizard Creek comprising the largest densities of redds of the four tributaries surveyed (22.9 redds/km).

<br>

PSCIS stream crossing `r as.character(my_site)` was ranked as a high priority for follow up with habitat confirmation due to the large size of the stream recorded in PSCIS (channel width = 2.4m) relative to other tributary streams in the Lizard Creek watershed and because it was rated as containing high value habitat by @vastFishPassage2013 and @grainger2011FishPassage2011. There is a bridge located on the Lazy Lizard recreation trail approximately 150m upstream of the crossing and an unassessed modelled crossing (ID 4600926) approximately 2100m upstream which could not be accessed due to a road closure.  The habitat confirmation was completed on `r my_pscis_info() %>% pull(date) %>% format(., format = "%B %d, %Y")`. A map of the watershed including areas surveyed is provided in Attachment 1 – Map [`r my_bcfishpass() %>% pull(map_tile_display_name)`](`r my_mapsheet()`).

<br>

## Stream Characteristics at Crossing {-}

As noted, the culvert located at PSCIS `r as.character(my_site)` appeared to have been replaced in 2020 with fresh rock, road fill and pavement present.  At the time of the survey, the un-embedded and non-backwatered `r my_pscis_info() %>% pull(diameter_or_span_meters)`m diameter crossing was considered a barrier to upstream fish passage with a pipe length of `r my_pscis_info() %>% pull(length_or_width_meters)`m, a culvert slope of `r my_pscis_info() %>% pull(culvert_slope_percent)`%, a stream width ratio of `r my_pscis_info() %>% pull(stream_width_ratio)` and an outlet drop of `r my_pscis_info() %>% pull(outlet_drop_meters)`m (Table \@ref(tab:tab-culvert-50152)). Water temperature was `r my_habitat_info() %>% filter(rowname == 'temperature c') %>% pull(us)`$^\circ$C, pH was `r my_habitat_info() %>% filter(rowname == 'p h') %>% pull(us)` and conductivity was `r my_habitat_info() %>% filter(rowname == 'conductivity m s cm') %>% pull(us)`uS/cm.  

<br>


```{r eval=F}
##this is useful to get some comments for the report
hab_site %>% filter(site == my_site & location == 'us') %>% pull(comments)
my_priority <-  my_priority_info()
my_hab_info <- my_habitat_info()

```

## Stream Characteristics Downstream {-}

The stream was surveyed downstream from the culvert for `r my_priority_info(loc = 'ds') %>% pull(survey_length_m)`m to Lizard Creek. Overall, total cover amount was rated as `r my_habitat_info2(loc = 'ds') %>% filter(rowname == 'total cover') %>% pull(v)` with  `r my_habitat_info2(loc = 'ds') %>% filter(v == 'dominant') %>% pull(rowname)` dominant. Cover was also present as `r my_habitat_info2(loc = 'ds') %>% filter(v == 'sub-dominant') %>% pull(rowname) %>% knitr::combine_words()` (Table \@ref(tab:tab-habitat-summary-50152), Figure \@ref(fig:photo-50152-01)). The average channel width was `r my_habitat_info3(loc = 'ds', row = 'avg channel width m')`m, the average wetted width was `r my_habitat_info3(loc = 'ds', row = 'avg wetted width m')`m and the average gradient was `r my_habitat_info3(loc = 'ds', row = 'average gradient percent')`%.  The dominant substrate was `r my_habitat_info3(loc = 'ds', row = "bed material dominant")` with `r my_habitat_info3(loc = 'ds', row = "bed material subdominant")` subdominant. Downstream of the crossing, numerous fry were observed, there were frequent sections of gravels suitable for salmonid spawning and there were no barriers or obstacles to fish passage. Habitat was rated as `r my_priority_info(loc = 'ds') %>%  pull(hab_value)` value for fry/juvenile salmonid rearing.  

<br>

## Stream Characteristics Upstream {-}

The stream was surveyed upstream from the culvert for `r my_priority_info(loc = 'us') %>% pull(survey_length_m)`m. Overall, total cover amount was rated as `r my_habitat_info2(loc = 'us') %>% filter(rowname == 'total cover') %>% pull(v)` with  `r my_habitat_info2(loc = 'us') %>% filter(v == 'dominant') %>% pull(rowname)` dominant. Cover was also present as `r my_habitat_info2(loc = 'us') %>% filter(v == 'sub-dominant') %>% pull(rowname) %>% knitr::combine_words()` (Table \@ref(tab:tab-habitat-summary-50152), Figure \@ref(fig:photo-50152-02)). The average channel width was `r my_habitat_info3(loc = 'us', row = 'avg channel width m')`m, the average wetted width was `r my_habitat_info3(loc = 'us', row = 'avg wetted width m')`m and the average gradient was `r my_habitat_info3(loc = 'us', row = 'average gradient percent')`%.  There were numerous fry observed throughout the area surveyed and abundant gravels suitable for resident westslope cutthrout trout spawning were noted. Habitat value was rated as `r my_priority_info(loc = 'us') %>% pull(hab_value)` for salmonid rearing and spawning. 


<br>

## Structure Remediation and Cost Estimate {-}

Structure replacement with an open bottomed structure is recommended to provide access to the habitat located upstream of PSCIS crossing `r as.character(my_site)`.  The cost of the work is estimated at \$`r as.character(my_cost_estimate() %>% pull(cost_est_1000s) * 1000)` for a cost benefit of \$`r as.character(my_cost_estimate() %>% pull(cost_net) * 1000)`/linear m and \$`r as.character(my_cost_estimate() %>% pull(cost_area_net) * 1000)`/m^2^.

<br>


## Conclusion {-}

There is `r my_priority_info() %>% pull(upstream_habitat_length_km)`km of mainstem habitat in Stove Creek upstream of crossing `r as.character(my_site)` with habitat rated as high value for fry and juvenile salmonid rearing. Of the four streams surveyed with habitat confirmations in the Lizard Creek drainage during this project, Stove Creek contained the highest value habitat. The large outlet drop (`r my_pscis_info() %>% pull(outlet_drop_meters)`m) on the newly installed culvert prevents upstream migration by all species and life stages.  This paved section of road is within Mt.Fernie Provincial Park. The crossing was ranked as a `r my_priority_info() %>% pull(priority)` priority for proceeding to design for replacement.

<br>

```{r tab-culvert-50152, eval = T}
print_tab_summary()
```

<br>

```{r tab-habitat-summary-50152}
tab_hab_summary %>% 
  filter(Site  == my_site) %>% 
  select(-Site) %>% 
  my_kable(caption_text = paste0('Summary of habitat details for PSCIS crossing ', my_site, '.'))

```


<br>

```{r photo-50152-01, fig.cap= my_caption, eval=T}
my_photo = pull_photo_by_str(str_to_pull = '_d1')

my_caption = paste0('Typical habitat downstream of PSCIS crossing ', my_site, '.')

grid::grid.raster(get_img())

# knitr::include_graphics(get_img_path())
```

<br>

```{r photo-50152-02, fig.cap= my_caption, eval=T}
my_photo = pull_photo_by_str(str_to_pull = '_u1')

my_caption = paste0('Typical habitat upstream of PSCIS crossing ', my_site, '.')

grid::grid.raster(get_img())
# knitr::include_graphics(get_img_path())
```


