---
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup-197542, eval = F}
knitr::opts_chunk$set(echo=FALSE, message=FALSE, warning=FALSE)
source('R/packages.R')
source('R/functions.R')
source('R/functions-phase2.R')
source('R/tables-phase2.R')
source('R/tables.R')
source('R/extract-fish.R')
source('R/functions-fish.R')


```



```{r  load-data-197542}
my_site <- 197542
my_site2 <- 197582
my_site_xref2 <- 4600157
```


```{r eval = F}
my_bcfish <- my_bcfishpass() ##can be useful to look at

# `r paste0('# Appendix - Crossing ', my_site)`{-} 
# `r paste0('**', my_overview_info() %>% pull(road_name), ' - ', my_overview_info() %>% pull(stream_name), '**')`
```

`r appendix_title()`
`r appendix_subtitle()`



## Site Location {-}

Crossing `r as.character(my_site)` is located on a `r my_pscis_info() %>% pull(stream_name)`, approximately `r my_priority_info(loc = 'ds') %>% pull(survey_length_m)`m upstream from the confluence with the Elk River. The crossing is located on `r my_overview_info() %>% pull(road_name)` just north of Fernie.  `r my_overview_info() %>% pull(road_name)` is a paved Ministry of Transportation and Infrastructure collector road accessing semi-rural residential and recreational areas. 


<br>

## Background {-}

At the crossing location, the stream is `r my_bcfishpass() %>% pull(stream_order) %>% english::ordinal()` order with a watershed area upstream of the road of approximately `r my_bcfishpass() %>% mutate(watershed_upstr_ha = watershed_upstr_ha/100) %>% pull(watershed_upstr_ha) %>% round(1)` km^2^. The elevation of the watershed ranges from a maximum of 2350 to 1030m at the culvert. The system is headed by the 1.6ha Hartley Lake. Although, there are several modelled crossings on the mainstem of Hartley Creek upstream of the crossing, 2020 surveys indicated they were either fords or bridges. PSCIS crossing 197582 is an open bottomed structure located under Highway 3 approximately 400m downstream of Dicken Road and was considered passable. 

<br>

During surveys, we spoke with the owner of lands adjacent to crossing `r as.character(my_site)` and they noted that the properties both upstream and downstream of the crossing were located within an area used historically as a log sort and that Hartley Creek was used to transport logs at that time.  The landowner also reported that within the last ten years there was a dam failure upstream near Hartley Lake that resulted in significant debris flows throughout the lower reaches of the stream. Aggradation of the channel adjacent to this crossing, reported by @robinson2008HartleyCreek as likely a result of a low width/depth ratio and unstable banks, has been an ongoing issue requiring repeated dredging to reduce disruptions to traffic flow and high water damage to highway infrastructure. Interior Reforestation Co. Ltd. designed and installed remediation measures in 2006 and 2007 (vortex weir installation, channel armouring, rootwad installation and riparian planting) with the intent to restore the hydraulic conditions necessary to carry transported gravel further downstream and improve fish habitat immediately below the crossing  [@robinson2008HartleyCreek]. The works were not completely successful in preventing aggradation, as dredging was reported in 2019 (Marlim Ecological Consulting 2019) with only approximately 30cm of headboard observed within the highway stream crossing structure during 2020 surveys.

<br>


Westslope cutthroat trout, bull trout and brook trout have been recorded in Hartley Creek upstream of PCSIS `r as.character(my_site)` with westslope cutthroat trout, bull trout and mountain whitefish recorded below [@data_fish_obs]. A radio-tagging study conducted in 2001–2002, identified Hartley Creek as a spawning location of westslope cutthrout trout captured and tagged in the Elk River [@schweigert_etal2017COSEWICassessment].

<br>

PSCIS stream crossing `r as.character(my_site)` was ranked as a high priority for follow up with habitat confirmation due to the large amount of potential habitat modelled upstream (13km) including the 1.6ha Hartley Lake. The habitat confirmation was completed on `r my_pscis_info() %>% pull(date) %>% format(., format = "%B %d, %Y")`. A map of the watershed including areas surveyed is provided in Attachment 1 – Map [`r my_bcfishpass() %>% pull(map_tile_display_name)`](`r my_mapsheet()`).

<br>

## Stream Characteristics at Crossing {-}

At the time of the survey, the un-embedded and non-backwatered `r my_pscis_info() %>% pull(diameter_or_span_meters)`m diameter oval crossing structure was considered a barrier to upstream fish passage with a pipe length of `r my_pscis_info() %>% pull(length_or_width_meters)`m, a culvert slope of `r my_pscis_info() %>% pull(culvert_slope_percent)`%, a stream width ratio of `r my_pscis_info() %>% pull(stream_width_ratio)` and an outlet drop of `r my_pscis_info() %>% pull(outlet_drop_meters)`m (Table \@ref(tab:tab-culvert-197542)). Water temperature was `r my_habitat_info() %>% filter(rowname == 'temperature c') %>% pull(us)`$^\circ$C, pH was `r my_habitat_info() %>% filter(rowname == 'p h') %>% pull(us)` and conductivity was `r my_habitat_info() %>% filter(rowname == 'conductivity m s cm') %>% pull(us)`uS/cm.  

<br>


```{r eval=F}
##this is useful to get some comments for the report
hab_site %>% filter(site == my_site & location == 'ds') %>% pull(comments)
hab_site %>% filter(site == my_site2 & location == 'ds') %>% pull(comments)
my_hab_info <- my_habitat_info()
```

## Stream Characteristics Downstream {-}

The stream was surveyed downstream from the culvert for 940m to the confluence with the Elk River including the 500m below PSCIS crossing `r as.character(my_site2)`. Within the 400m section between Dicken Road and Highway 3, total cover amount was rated as `r my_habitat_info2(loc = 'ds') %>% filter(rowname == 'total cover') %>% pull(v)` with  `r my_habitat_info2(loc = 'ds') %>% filter(v == 'dominant') %>% pull(rowname)` dominant. Cover was also present as `r my_habitat_info2(loc = 'ds') %>% filter(v == 'sub-dominant') %>% pull(rowname) %>% knitr::combine_words()` (Table \@ref(tab:tab-habitat-summary-197542), Figure \@ref(fig:photo-197542-01)). The average channel width was `r my_habitat_info3(loc = 'ds', row = 'avg channel width m')`m, the average wetted width was `r my_habitat_info3(loc = 'ds', row = 'avg wetted width m')`m and the average gradient was `r my_habitat_info3(loc = 'ds', row = 'average gradient percent')`%.  The dominant substrate was `r my_habitat_info3(loc = 'ds', row = "bed material dominant")` with `r my_habitat_info3(loc = 'ds', row = "bed material subdominant")` subdominant.  Although abundant gravels suitable for spawning were present, the Hartley Creek channel appeared anthropogenically straightened in the area surveyed with a notable lack of deep pools and undercut banks. Overall, habitat was rated as `r my_priority_info(loc = 'ds') %>%  pull(hab_value)` value providing suitable habitat for resident and fluvial westlope cutthrout trout spawning.

<br>

Within the 500m below PSCIS crossing `r as.character(my_site2)`, the first 175m of channel downstream of the culvert was primarily composed of an aggrading gravel channel with widths between 12-15m.  Downstream of this initial section was a marsh area impounded by multiple beaver dams containing run sections to 1m deep and fine substrates.  Much of the channel surveyed within the lower sections was within the Elk River floodplain.   Total cover amount was rated as `r my_habitat_info2(loc = 'ds', sit = my_site2) %>% filter(rowname == 'total cover') %>% pull(v)` with  `r my_habitat_info2(sit = my_site2, loc = 'ds') %>% filter(v == 'dominant') %>% pull(rowname)` dominant. Cover was also present as `r my_habitat_info2(sit = my_site2, loc = 'ds') %>% filter(v == 'sub-dominant') %>% pull(rowname) %>% knitr::combine_words()` (Table \@ref(tab:tab-habitat-summary-197542), Figures \@ref(fig:photo-197542-02) - \@ref(fig:photo-197542-03)). The average channel width was `r my_habitat_info3(sit = my_site2, loc = 'ds', row = 'avg channel width m')`m, the average wetted width was `r my_habitat_info3(sit = my_site2, loc = 'ds', row = 'avg wetted width m')`m and the average gradient was `r my_habitat_info3(sit = my_site2, loc = 'ds', row = 'average gradient percent')`%.  The dominant substrate was `r my_habitat_info3(sit = my_site2, loc = 'ds', row = "bed material dominant")` with `r my_habitat_info3(sit = my_site2, loc = 'ds', row = "bed material subdominant")` subdominant. Young of the year salmonids were observed upstream of upper most beaver dam. Habitat was rated as `r my_priority_info(sit = my_site2, loc = 'ds') %>%  pull(hab_value)` representing an important migration corridor containing habitat with moderate rearing potential for resident and fluvial salmonids.


```{r eval=F}
##this is useful to get some comments for the report
hab_site %>% filter(site == my_site & location == 'us') %>% pull(comments)
my_hab_info <- my_habitat_info()
```


<br>

## Stream Characteristics Upstream {-}

The stream was surveyed immediately upstream from the culvert for 650m with an addition 75m section surveyed upstram approximatley 2.5km upstream. A small dam (0.3m) high is located approximately 5m upstream of Dicken Road and may present an obstruction to upstream fish passage for fry and juvenile salmonids at some flows.  Overall, total cover amount was rated as `r my_habitat_info2(loc = 'us') %>% filter(rowname == 'total cover') %>% pull(v)` with  `r my_habitat_info2(loc = 'us') %>% filter(v == 'dominant') %>% pull(rowname)` dominant. Cover was also present as `r my_habitat_info2(loc = 'us') %>% filter(v == 'sub-dominant') %>% pull(rowname) %>% knitr::combine_words()` (Table \@ref(tab:tab-habitat-summary-197542), Figure \@ref(fig:photo-197542-02)). The average channel width was `r my_habitat_info3(loc = 'us', row = 'avg channel width m')`m, the average wetted width was `r my_habitat_info3(loc = 'us', row = 'avg wetted width m')`m and the average gradient was `r my_habitat_info3(loc = 'us', row = 'average gradient percent')`%.  The dominant substrate was `r my_habitat_info3(loc = 'us', row = "bed material dominant")` with `r my_habitat_info3(loc = 'us', row = "bed material subdominant")` subdominant. There were frequent pools formed by small and large woody debris ranging from 0.3 - 0.7m in depth with a debris jam near top of site (11U 643668 5491243) measured at 1.2m in height.  Pockets of gravels suitable for resident and fluvial salmonids were present and there were areas of erosion notable on the right bank of the stream. 

<br>

Within the 75m section of stream surveyed 2.5km upstream of Dicken Road, stream conditions were similiar to those observed downstream with an average channel width of 4.6m and gravel dominated substrate (Figure \@ref(fig:photo-197542-02b)). Overall, upstream of PSCIS `r as.character(my_site)` was rated as `r my_priority_info(loc = 'us') %>% pull(hab_value)` habitat value, containing habitat with moderate rearing and spawning potential for resident and fluvial salmonids.  

<br>


## Structure Remediation and Cost Estimate {-}

Structure replacement with an open bottomed structure is recommended to provide access to the habitat located upstream of PSCIS crossing `r as.character(my_site)`.  The cost of the work is estimated at \$`r as.character(my_cost_estimate() %>% pull(cost_est_1000s) * 1000)` for a cost benefit of \$`r as.character(my_cost_estimate() %>% pull(cost_net) * 1000)`/linear m and \$`r as.character(my_cost_estimate() %>% pull(cost_area_net) * 1000)`/m^2^.

<br>

## Conclusion {-}

There is an estimated `r my_priority_info() %>% pull(upstream_habitat_length_km)`km of mainstem habitat available upstream of crossing `r as.character(my_site)` with habitat rated as `r my_priority_info(loc = 'us') %>% pull(hab_value)` value, representing an important migration corridor with moderate rearing and spawning potential for resident and fluvial salmonids. `r my_overview_info() %>% pull(road_name)` is the responsiblity of the Ministry of Transportation and Infrastructure. The crossing was ranked as a `r my_priority_info() %>% pull(priority)` priority for proceeding to design for replacement with an open bottomed structure. The private land adjacent to Hartley Creek in the vicinity of Dicken Road and Highway 3 may represent valuable opportunities for channel and riparian restoration activities as it is located within an area of intensive fish habitat impacts resulting from dredging to protect highway infrastructure  as well as other historic land use activities (i.e. log transport and processing). 

\pagebreak

```{r tab-culvert-197542, eval = T}
print_tab_summary()

```

\pagebreak

```{r tab-habitat-summary-197542}
tab_hab_summary %>% 
  filter(Site  == my_site | Site  == my_site2) %>% 
  # select(-Site) %>% 
  my_kable(caption_text = paste0('Summary of habitat details for PSCIS crossing ', my_site, '.'))

```



<br>

```{r photo-197542-01, fig.cap= my_caption, out.width = photo_width, eval=T}
my_photo = pull_photo_by_str(str_to_pull = '_d1_')

my_caption = paste0('Typical habitat downstream of PSCIS crossing ', my_site, '.')

grid::grid.raster(get_img())

# knitr::include_graphics(get_img_path())
```

<br>

```{r photo-197542-01b, fig.cap= my_caption, out.width = photo_width, eval=T}
my_photo = pull_photo_by_str(str_to_pull = '_d2_')

my_caption = paste0('Typical habitat downstream of PSCIS crossing ', my_site2, '.')

grid::grid.raster(get_img())

# knitr::include_graphics(get_img_path())
```

<br>

```{r photo-197542-01c, fig.cap= my_caption, out.width = photo_width, eval=T}
my_photo = pull_photo_by_str(site_id = my_site_xref2, str_to_pull = 'crossing_')

my_caption = paste0('PSCIS crossing ', my_site2, '.')

grid::grid.raster(get_img(site = my_site_xref2))

# knitr::include_graphics(get_img_path())
```

<br>


```{r photo-197542-02, fig.cap= my_caption, out.width = photo_width, eval=T}

# 
my_photo = pull_photo_by_str(str_to_pull = '_u1_')

my_caption = paste0('Typical habitat upstream of PSCIS crossing ', my_site, '.')

grid::grid.raster(get_img())
# knitr::include_graphics(get_img_path())
```

<br>

```{r photo-197542-02b, fig.cap= my_caption, out.width = photo_width, eval=T}

# 
my_photo = pull_photo_by_str(str_to_pull = '_u2_')

my_caption = paste0('Habitat upstream of PSCIS crossing ', my_site, ' approximatley 2.5km.')

grid::grid.raster(get_img())
# knitr::include_graphics(get_img_path())
```



