
```{r setup-62423, eval = F}
knitr::opts_chunk$set(echo=FALSE, message=FALSE, warning=FALSE)
source('R/packages.R')
source('R/functions.R')
source('R/functions-phase2.R')
source('R/tables-phase2.R')
source('R/tables.R')
source('R/extract-fish.R')
source('R/functions-fish.R')


```



```{r  load-data-62423}
my_site <- 62423
```


`r paste0('# Appendix - Crossing ', my_site)`{-} 
`r paste0('**', my_overview_info() %>% pull(road_name), ' - ', my_overview_info() %>% pull(stream_name), '**')`


## Site Location {-}

Crossing `r as.character(my_site)` is located on `r my_pscis_info() %>% pull(stream_name)`, approximately `r my_priority_info(loc = 'ds') %>% pull(survey_length_m)`m upstream from the confluence with Grave Creek and accessed from Grave Creek Forest Service Road.  Although the freshwater atlas stream layer mapping incorrectly indicates the subject stream is the mainstem of Grave Creek it is actually a small drainage that joins Harriet Lake Creek which flows primarily from the 6ha Harriet Lake located to the north at an elevation of 2100m approximately 4km upstream of the crossing.

<br>

## Background {-}

Harriet Lake was stocked with "wild" westslope cutthrout trout five times between 1985 and 2002 [@bcmoeFishInventoriesData2020].  Downstream, Grave Creek is known to contain westslope cutthrout trout, rainbow trout and bull trout [@data_fish_obs]. Two habitat confirmation assessments were conducted downstream on the mainstem of Grave Creek in 2014 at PSCIS crossings 62421 and 62422 [@masseEKConfirmation2015]. Although Heather Lamson - MoE Fisheries Biologist recommended the culverts not be removed to prevent potential hybridization of westslope cutthrout trout with stocked rainbow trout downstream [@masseEKConfirmation2015], both structures had been replaced with bridges at the time of the surveys. Designs and remediations of these crossings were not recorded in PSCIS. Reassessments of these crossings were conducted by our team in 2020 and results will be loaded to PSCIS. In the field, Lotic Environmental Ltd. field teams were observed conducting two-pass closed site electrofishing in Grave Creek as part of a westslope cutthrout trout population assessment and aquatic monitoring program. Data from the program is uploaded to the Fisheries Information Summary System annually and is made available through the BC Data Catalog [@data_fish_obs; @moeStreamInventorySample].

<br>

 
Although the modelling of potential habitat upstream of this crossing was considered not accurate due to the incorrect mapping of Grave Creek, during field work planning, PSCIS stream crossing `r as.character(my_site)` was ranked as a high priority for follow up with habitat confirmation due to the large size of the modelled stream network upstream (20km) and because it was located on a stream with habitat rated as moderate value by @vastFishPassage2013. A bridge (PSCIS 62413) is located approximately 1km upstream of the crossing. The habitat confirmation was completed on `r my_pscis_info() %>% pull(date) %>% format(., format = "%B %d, %Y")`.  A map of the watershed including areas surveyed is provided in Attachment 1 – Map [`r my_bcfishpass() %>% pull(map_tile_display_name)`](`r my_mapsheet()`).

<br>


## Stream Characteristics at Crossing {-}

At the time of the survey, the un-embedded and non-backwatered `r my_pscis_info() %>% pull(diameter_or_span_meters)`m diameter crossing was considered a barrier to upstream fish passage with a pipe length of `r my_pscis_info() %>% pull(length_or_width_meters)`m, a culvert slope of `r my_pscis_info() %>% pull(culvert_slope_percent)`%, a stream width ratio of `r my_pscis_info() %>% pull(stream_width_ratio)` and an outlet drop of `r my_pscis_info() %>% pull(outlet_drop_meters)`m (Table \@ref(tab:tab-culvert-62423)). The stream was dry at the crossing location at the time of the survey. 

<br>


```{r eval=F}
##this is useful to get some comments for the report
hab_site %>% filter(site == my_site & location == 'us') %>% pull(comments)

```
## Stream Characteristics Downstream {-}

The stream was surveyed downstream from the culvert for `r my_priority_info(loc = 'ds') %>% pull(survey_length_m)`m to Grave Creek. Overall, total cover amount was rated as `r my_habitat_info2(loc = 'ds') %>% filter(rowname == 'total cover') %>% pull(v)` with  `r my_habitat_info2(loc = 'ds') %>% filter(v == 'dominant') %>% pull(rowname)` dominant. Cover was also present as `r my_habitat_info2(loc = 'ds') %>% filter(v == 'sub-dominant') %>% pull(rowname) %>% knitr::combine_words()` (Table \@ref(tab:tab-habitat-summary-62423), Figure \@ref(fig:photo-62423-01)). The average channel width was `r my_habitat_info3(loc = 'ds', row = 'avg channel width m')`m, and the average gradient was `r my_habitat_info3(loc = 'ds', row = 'average gradient percent')`%.  The dominant substrate was `r my_habitat_info3(loc = 'ds', row = "bed material dominant")` with `r my_habitat_info3(loc = 'ds', row = "bed material subdominant")` subdominant. Habitat value was rated as `r my_priority_info(loc = 'ds') %>% pull(hab_value)` value due to a lack of flow.  

<br>

## Stream Characteristics Upstream {-}

The stream was surveyed upstream from the culvert in the general location of the mapped Grave Creek streamline for `r my_priority_info(loc = 'us') %>% pull(survey_length_m)`m. As Harriet Lake Creek was the primary source of flow for the area and enters the surveyed tributary 150m upstream of the crossing location, flows were very minimal and substrate was primarily fines above its confluence. There was however, a visible channel in this location with a ford (PSCIS 197563) located approximately 600m upstream fo PSCSIS `r as.character(my_site)`. Overall, total cover amount was rated as `r my_habitat_info2(loc = 'us') %>% filter(rowname == 'total cover') %>% pull(v)` with  `r my_habitat_info2(loc = 'us') %>% filter(v == 'dominant') %>% pull(rowname)` dominant. Cover was also present as `r my_habitat_info2(loc = 'us') %>% filter(v == 'sub-dominant') %>% pull(rowname) %>% knitr::combine_words()` (Table \@ref(tab:tab-habitat-summary-62423), Figure \@ref(fig:photo-62423-02)). The average channel width was `r my_habitat_info3(loc = 'us', row = 'avg channel width m')`m, the average wetted width was `r my_habitat_info3(loc = 'us', row = 'avg wetted width m')`m and the average gradient was `r my_habitat_info3(loc = 'us', row = 'average gradient percent')`%.  Harriet Lake Creek gradients modelled at 19% just upstream of the bridge located 1km upstream and too steep for upstream westslope cutthrout passage (>20%) at 2.2km. Habitat value was rated as `r my_priority_info(loc = 'us') %>% pull(hab_value)` for fry and juvenile westslope cutthrout rearing in Harriet Creek and low in the unnamed tributary mapped as Grave Creek due to a lack of flow. 


<br>

## Structure Remediation and Cost Estimate {-}

Structure replacement with an embedded culvert is recommended to provide access to the habitat located upstream of PSCIS crossing `r as.character(my_site)`.  The cost for the work is estimated at \$`r as.character(my_cost_estimate() %>% pull(cost_est_1000s) * 1000)` for a cost benefit of \$`r as.character(my_cost_estimate() %>% pull(cost_net) * 1000)`/linear m and \$`r as.character(my_cost_estimate() %>% pull(cost_area_net) * 1000)`/m^2^.


## Conclusion {-}

There is `r my_priority_info() %>% pull(upstream_habitat_length_km)`km of mainstem habitat upstream of crossing `r as.character(my_site)` with habitat in the areas surveyed upstream of the crossing rated as `r my_priority_info(loc = 'us') %>% pull(hab_value)` value. Although the provincial forest tenure road layer does not include Grave Creek FSR, it is likely that it is a tenure road under the responsibility of the Ministry of Forests, Lands, Natural Resource Operations and Rural Development.  The crossing was ranked as a `r my_priority_info() %>% pull(priority)` priority for proceeding to design for replacement with an open bottomed structure due to the low water conditions and steep gradients upstream in Harriet Lake Creek.

<br>

```{r tab-culvert-62423, eval = T}
##had to pull this object out of the table push for some reason
print_tab_summary()

```

<br>

```{r tab-habitat-summary-62423}
tab_hab_summary %>% 
  filter(Site  == my_site) %>% 
  select(-Site) %>% 
  my_kable(caption_text = paste0('Summary of habitat details for PSCIS crossing ', my_site, '.'))

```


<br>

```{r photo-62423-02, fig.cap= my_caption, out.width = photo_width, eval=T}
my_photo = '20200920_110902_k_d01_.JPG'

my_caption = paste0('Typical habitat downstream of PSCIS crossing ', my_site, '.')

grid::grid.raster(get_img())

# knitr::include_graphics(get_img_path())
```

<br>

```{r photo-62423-01, fig.cap= my_caption, out.width = photo_width, eval=T}

# 
my_photo = '20200920_113417_k_u01_.JPG'

my_caption = paste0('Typical habitat upstream of PSCIS crossing ', my_site, '.')

grid::grid.raster(get_img())
# knitr::include_graphics(get_img_path())
```

