
```{r setup-01, eval = F}
knitr::opts_chunk$set(echo=FALSE, message=FALSE, warning=FALSE)
source('R/packages.R')
source('R/functions.R')
source('R/functions-phase2.R')
source('R/tables-phase2.R')
source('R/tables.R')
source('R/extract-fish.R')
source('R/functions-fish.R')


```



```{r  load-data-01}
my_site <- 50155
```


`r paste0('# Appendix - Crossing ', my_site)`{-} 
`r paste0('**', my_overview_info() %>% pull(road_name), ' - ', my_overview_info() %>% pull(stream_name), '**')`


## Site Location {-}

Crossing `r as.character(my_site)` is located on a `r my_pscis_info() %>% pull(stream_name)`, approximately 75m upstream from the confluence with Lizard Creek. The stream is located approximately 100m east of the location where it is mapped on the freshwater atlas stream layer. Island Lake Lodge Road is an extension of Mt.Fernie Park Road with access to Highway 3 located within Fernie city limits.  The are is a popular recreational destination for hikers and mountain bikers.  Island Lake Lodge is a year round tourist destination providing accomodations, guided hiking and backcountry catskiing for clients.  


<br>

## Background {-}

At the crossing location, the stream is 2nd order with a watershed area upstream of the road of approximately 1.8km^2^. The elevation of the watershed ranges from a maximum of 1945m to 1080m at the culvert. One 12m long bridge (PSCIS 197543) is located upstream of the subject crossing approximately 575m on the Lazy Lizard bike trail and another 7m long bridge structure is located downstream also on a recreational trail. A search of provincial records yielded no fisheries information for the stream [@moeStreamInventorySample]. Downstream, Lizard Creek supports westslope cutthroat trout, bull trout, mountain whitefish, brook trout, longnose sucker and longnose dace [@data_fish_obs]. @elkriveralliance2020ElkRiver conducted redd surveys in Lizard Creek in 2019 with a total of 55 redds observed within a 2.4km of stream surveyed comprising the largest densities of redds of the four tributaries surveyed.

<br>

PSCIS stream crossing `r as.character(my_site)` was ranked as a high priority for follow up with habitat confirmation due to the large size of the stream relative to other tributary streams in the watershed, the previously rated high value habitat and because it was prioritized for follow up by @vastFishPassage2013. The habitat confirmation was completed on `r my_pscis_info() %>% pull(date) %>% format(., format = "%B %d, %Y")`. A map of the watershed including areas surveyed is provided in Attachment 1 – Map [`r my_bcfishpass() %>% pull(map_tile_display_name)`](`r my_mapsheet()`).

<br>

## Stream Characteristics at Crossing {-}

At the time of the survey, the un-embedded and non-backwatered `r my_pscis_info() %>% pull(diameter_or_span_meters)`m diameter crossing was considered a barrier to upstream fish passage with a pipe length of `r my_pscis_info() %>% pull(length_or_width_meters)`m, a culvert slope of `r my_pscis_info() %>% pull(culvert_slope_percent)`%, a stream width ratio of `r my_pscis_info() %>% pull(stream_width_ratio)` an outlet drop of `r my_pscis_info() %>% pull(outlet_drop_meters)`m (Table \@ref(tab:tab-culvert-010)). Water temperature was `r my_habitat_info() %>% filter(rowname == 'temperature c') %>% pull(us)`$^\circ$C, pH was `r my_habitat_info() %>% filter(rowname == 'p h') %>% pull(us)` and conductivity was `r my_habitat_info() %>% filter(rowname == 'conductivity m s cm') %>% pull(us)`uS/cm.  

<br>


```{r eval=F}
##this is useful to get some comments for the report
hab_site %>% filter(site == my_site & location == 'us') %>% pull(comments)

```
## Stream Characteristics Downstream {-}

The stream was surveyed downstream from the culvert for `r my_priority_info(loc = 'ds') %>% pull(survey_length_m)`m to Lizard Creek. Overall, total cover amount was rated as `r my_habitat_info2(loc = 'ds') %>% filter(rowname == 'total cover') %>% pull(v)` with  `r my_habitat_info2(loc = 'ds') %>% filter(v == 'dominant') %>% pull(rowname)` dominant. Cover was also present as `r my_habitat_info2(loc = 'ds') %>% filter(v == 'sub-dominant') %>% pull(rowname) %>% knitr::combine_words()` (Table \@ref(tab:tab-habitat-summary-020), Figure \@ref(fig:photo-001)). The average channel width was `r my_habitat_info3(loc = 'ds', row = 'avg channel width m')`m, the average wetted width was `r my_habitat_info3(loc = 'ds', row = 'avg wetted width m')`m and the average gradient was `r my_habitat_info3(loc = 'ds', row = 'average gradient percent')`%.  Habitat value was rated as medium with good potential for fry/juvenile salmonid rearing but a lack of deep pools for adult overwintering and rearing. 

<br>

## Stream Characteristics Upstream {-}

The stream was surveyed upstream from the culvert for `r my_priority_info(loc = 'us') %>% pull(survey_length_m)`m. Overall, total cover amount was rated as `r my_habitat_info2(loc = 'us') %>% filter(rowname == 'total cover') %>% pull(v)` with  `r my_habitat_info2(loc = 'us') %>% filter(v == 'dominant') %>% pull(rowname)` dominant. Cover was also present as `r my_habitat_info2(loc = 'us') %>% filter(v == 'sub-dominant') %>% pull(rowname) %>% knitr::combine_words()` (Table \@ref(tab:tab-habitat-summary-020), Figure \@ref(fig:photo-002)). The average channel width was `r my_habitat_info3(loc = 'us', row = 'avg channel width m')`m, the average wetted width was `r my_habitat_info3(loc = 'us', row = 'avg wetted width m')`m and the average gradient was `r my_habitat_info3(loc = 'us', row = 'average gradient percent')`%.  There were frequent areas of gravels suitable for resident westslope cutthrout trout spawning.  Frequent pools to 40cm deep were present and associated with small and large woody debris. Habitat value was rated as `r my_priority_info(loc = 'us') %>% pull(hab_value)` for fry and juvenile westslope cutthrout rearing. 

<br>

## Fish Sampling {-}

To assess potential impacts of the culvert on fish densities in the stream we electrofished upstream and downstream of the crossing.  Three sites were sampled upstream and one site  was sampled downstream. A total of `r tab_fish_summary %>% filter(site_id == paste0(my_site, '_us') & species_code == 'WCT') %>% pull(count_fish)` westslope cutthrout trout and `r tab_fish_summary %>% filter(site_id == paste0(my_site, '_us') & species_code == 'EB') %>% pull(count_fish)` eastern brook trout were captured upstream with `r tab_fish_summary %>% filter(site_id == paste0(my_site, '_ds') & species_code == 'WCT') %>% pull(count_fish)` westslope cutthrout trout and `r tab_fish_summary %>% filter(site_id == paste0(my_site, '_ds') & species_code == 'EB') %>% pull(count_fish)` eastern brook trout captured downstream.  Raw results are included in digital format as [Attachment 2](https://github.com/NewGraphEnvironment/fish_passage_elk_2020_reporting/raw/master/data/habitat_confirmations.xls) and summarized in Tables \@ref(tab:tab-fish-site-010) - \@ref(tab:tab-fish-dens-010) and Figure \@ref(fig:plot-fish-box-010).

<br>

## Structure Remediation and Cost Estimate {-}

Structure replacement with an open bottomed culvert is recommended to provide unconstained access to the habitat located upstream of PSCIS crossing `r as.character(my_site)`.  The cost for work is estimated at \$`r as.character(my_cost_estimate() %>% pull(cost_est_1000s) * 1000)` for a cost benefit of \$`r as.character(my_cost_estimate() %>% pull(cost_net) * 1000)`/linear m and \$`r as.character(my_cost_estimate() %>% pull(cost_area_net) * 1000)`/m^2^.


## Conclusion {-}

There is an estimated `r my_priority_info() %>% pull(upstream_habitat_length_km)`km of mainstem habitat upstream of crossing `r as.character(my_site)` with habitat in the areas surveyed upstream of the crossing rated as high value for fry and juvenile salmonid rearing. Although potentially attributable to the lower gradient habitat downstream and the proximity to the Lizard Creek mainstem, fish sampling results indicated that the crossing is potentially negatively impacting habitat capacity upstream of the crossing as higher densities of westslope cutthrout trout were captured downstream of the crossing than above. The road may be part of the Island Lake Recreational tenure or solely the responsiblity of the Ministry of Forests, Lands, Natural Resource Operations & Rural Development. The crossing was ranked as a `r my_priority_info() %>% pull(priority)` priority for proceeding to design for replacement with an open bottomed structure.

\pagebreak

```{r tab-culvert-010}
print_tab_summary()


```

\pagebreak

```{r tab-habitat-summary-010}
tab_hab_summary %>% 
  filter(Site  == my_site) %>% 
  select(-Site) %>% 
  my_kable(caption_text = paste0('Summary of habitat details for PSCIS crossing ', my_site, '.'))

```

<br>

```{r tab-fish-site-010, eval=T}
my_caption <- paste0('Electrofishing sites for PSCIS crossing ', my_site, '.')

tab_fish_site() %>% 
  my_kable(caption_text = my_caption)
```

<br>

```{r tab-fish-dens-010, eval=T}
my_caption <- paste0('Westslope cuthrout trout densities (fish/100m2) for PSCIS crossing ', my_site, '.')

tab_fish_dens() %>% 
  my_kable(caption_text = my_caption)
```

<br>

```{r plot-fish-box-010, fig.cap= my_caption, eval=T}
my_caption <- paste0('Fish densities (fish/100m2) for PSCIS crossing ', my_site, '.')

p <- plot_fish_box() 
p
```


<br>

```{r photo-001, fig.cap= my_caption, eval=T}
my_photo = '20200922_131045_k.JPG'

my_caption = paste0('Typical habitat downstream of PSCIS crossing ', my_site, '.')

grid::grid.raster(get_img())

# knitr::include_graphics(get_img_path())
```

<br>

```{r photo-002, fig.cap= my_caption, eval=T}

# 
my_photo = 'IMG_3304_k.JPG'

my_caption = paste0('Typical habitat upstream of PSCIS crossing ', my_site, '.')

grid::grid.raster(get_img())
# knitr::include_graphics(get_img_path())
```

