---
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup-50181, eval = F}
knitr::opts_chunk$set(echo=FALSE, message=FALSE, warning=FALSE)
source('R/packages.R')
source('R/functions.R')
source('R/functions-phase2.R')
source('R/tables-phase2.R')
source('R/tables.R')
source('R/extract-fish.R')
source('R/functions-fish.R')


```



```{r  load-data-50181}
my_site <- 50181

```


```{r eval = F}
my_bcfish <- my_bcfishpass() ##can be useful to look at

# `r paste0('# Appendix - Crossing ', my_site)`{-} 
# `r paste0('**', my_overview_info() %>% pull(road_name), ' - ', my_overview_info() %>% pull(stream_name), '**')`
```

`r appendix_title()`
`r appendix_subtitle()`



## Site Location {-}

Crossing `r as.character(my_site)` is located on a `r my_pscis_info() %>% pull(stream_name)`, approximately `r my_priority_info(loc = 'ds') %>% pull(survey_length_m)`m upstream from the confluence with a tributary to Morrissey Creek. The crossing is located on `r my_overview_info() %>% pull(road_name)` just south of Morrissey approximately 15km south of Fernie.  `r my_overview_info() %>% pull(road_name)` is a gravel forest tenure road with active log hauling at the time of the survey. This crossing is on a stream that flows into the Morrissey Creek tributary approximately 2.7km upstream of where PSCIS crossing 50185 on River Road was also surveyed with a habitat confirmation assessment ([Appendix - Crossing 50185]).  


<br>

## Background {-}

At the crossing location, the stream is `r my_bcfishpass() %>% pull(stream_order) %>% english::ordinal()` order with a watershed area upstream of the road of approximately `r my_bcfishpass() %>% mutate(watershed_upstr_ha = watershed_upstr_ha/100) %>% pull(watershed_upstr_ha) %>% round(1)` km^2^. The elevation of the watershed ranges from a maximum of 2100 to 1200m at the culvert. Upstream of the crossing, although there are four fords documented in PSCIS there are no other road-stream crossing barriers. No fisheries information was available for the stream [@moeStreamInventorySample] however westslope cutthroat trout, bull trout, mountain whitefish, brook trout, longnose sucker and longnose dace have been recorded downstream in Morrissey Creek [@data_fish_obs]. @elkriveralliance2020ElkRiver conducted redd surveys in Morrissey Creek in 2019 with a total of 7 redds observed within a 3.6km of stream surveyed.  

<br>

PSCIS stream crossing `r as.character(my_site)` was ranked as a high priority for follow up with habitat confirmation due to the relatively large channel width (3.7m) and the previously rated high value habitat reported in PSCIS from @grainger2011FishPassage2011. The habitat confirmation was completed on `r my_pscis_info() %>% pull(date) %>% format(., format = "%B %d, %Y")`. A map of the watershed including areas surveyed is provided in Attachment 1 – Map [`r my_bcfishpass() %>% pull(map_tile_display_name)`](`r my_mapsheet()`).

<br>

## Stream Characteristics at Crossing {-}

At the time of the survey, the un-embedded and non-backwatered `r my_pscis_info() %>% pull(diameter_or_span_meters)`m diameter crossing was considered a barrier to upstream fish passage with a pipe length of `r my_pscis_info() %>% pull(length_or_width_meters)`m, a culvert slope of `r my_pscis_info() %>% pull(culvert_slope_percent)`%, a stream width ratio of `r my_pscis_info() %>% pull(stream_width_ratio)` and an outlet drop of `r my_pscis_info() %>% pull(outlet_drop_meters)`m (Table \@ref(tab:tab-culvert-50181)). Water temperature was `r my_habitat_info() %>% filter(rowname == 'temperature c') %>% pull(us)`$^\circ$C, pH was `r my_habitat_info() %>% filter(rowname == 'p h') %>% pull(us)` and conductivity was `r my_habitat_info() %>% filter(rowname == 'conductivity m s cm') %>% pull(us)`uS/cm.  

<br>


```{r eval=F}
##this is useful to get some comments for the report
hab_site %>% filter(site == my_site & location == 'ds') %>% pull(comments)
my_hab_info <- my_habitat_info()
```

## Stream Characteristics Downstream {-}

The stream was surveyed downstream from the culvert for `r my_priority_info(loc = 'ds') %>% pull(survey_length_m)`m to the confluence with the downstream channel. The mouth of the stream is located approximately 100m east of where it is mapped on the freshwater atlas stream layer. Overall, total cover amount was rated as `r my_habitat_info2(loc = 'ds') %>% filter(rowname == 'total cover') %>% pull(v)` with  `r my_habitat_info2(loc = 'ds') %>% filter(v == 'dominant') %>% pull(rowname)` dominant. Cover was also present as `r my_habitat_info2(loc = 'ds') %>% filter(v == 'sub-dominant') %>% pull(rowname) %>% knitr::combine_words()` (Table \@ref(tab:tab-habitat-summary-50181), Figure \@ref(fig:photo-50181-01)). The average channel width was `r my_habitat_info3(loc = 'ds', row = 'avg channel width m')`m, the average wetted width was `r my_habitat_info3(loc = 'ds', row = 'avg wetted width m')`m and the average gradient was `r my_habitat_info3(loc = 'ds', row = 'average gradient percent')`%.  The dominant substrate was `r my_habitat_info3(loc = 'ds', row = "bed material dominant")` with `r my_habitat_info3(loc = 'ds', row = "bed material subdominant")` subdominant.  Intermittent smaller pools and pockets of gravels suitably sized for resident salmonid spawning were present. Habitat was rated as `r my_priority_info(loc = 'ds') %>%  pull(hab_value)` value for fry/juvenile salmonid rearing.


```{r eval=F}
##this is useful to get some comments for the report
hab_site %>% filter(site == my_site & location == 'us') %>% pull(comments)
my_hab_info <- my_habitat_info()
```


<br>

## Stream Characteristics Upstream {-}

The stream was surveyed upstream from the culvert for `r my_priority_info(loc = 'us') %>% pull(survey_length_m)`m. Overall, total cover amount was rated as `r my_habitat_info2(loc = 'us') %>% filter(rowname == 'total cover') %>% pull(v)` with  `r my_habitat_info2(loc = 'us') %>% filter(v == 'dominant') %>% pull(rowname)` dominant. Cover was also present as `r my_habitat_info2(loc = 'us') %>% filter(v == 'sub-dominant') %>% pull(rowname) %>% knitr::combine_words()` (Table \@ref(tab:tab-habitat-summary-50181), Figure \@ref(fig:photo-50181-02)). The average channel width was `r my_habitat_info3(loc = 'us', row = 'avg channel width m')`m, the average wetted width was `r my_habitat_info3(loc = 'us', row = 'avg wetted width m')`m and the average gradient was `r my_habitat_info3(loc = 'us', row = 'average gradient percent')`%.  The dominant substrate was `r my_habitat_info3(loc = 'us', row = "bed material dominant")` with `r my_habitat_info3(loc = 'us', row = "bed material subdominant")` subdominant. Good flows were present and some pockets of gravels suitable for spawning were observed throughout.  There were steps ranging from 0.5 - 0.8m due to large woody debris jams but no permanent migration barriers were observed until a cascade at the top end of the survey area (24% for >50m). Habitat was rated as `r my_priority_info(loc = 'us') %>% pull(hab_value)` value for fry and juvenile westslope cutthrout trout rearing. 

<br>

## Fish Sampling {-}

To confirm fish presence and assess potential impacts of the culvert on fish densities in the stream, electrofishing was conducted upstream and downstream of the crossing.  One site was sampled downstream and one site was sampled upstream. A total of `r tab_fish_summary %>% filter(site_id == paste0(my_site, '_us') & species_code == 'WCT') %>% pull(count_fish)` westslope cutthrout trout were captured upstream and `r tab_fish_summary %>% filter(site_id == paste0(my_site, '_ds') & species_code == 'WCT') %>% pull(count_fish)` were captured downstream (Figure \@ref(fig:photo-50181-03)).  Raw results are included in digital format as [Attachment 2](https://github.com/NewGraphEnvironment/fish_passage_elk_2020_reporting/raw/master/data/habitat_confirmations.xls) and summarized in Tables \@ref(tab:tab-fish-site-50181) - \@ref(tab:tab-fish-dens-50181) and Figure \@ref(fig:plot-fish-box-50181).

<br>

## Structure Remediation and Cost Estimate {-}

Structure replacement with an open bottomed structure is recommended to provide access to the habitat located upstream of PSCIS crossing `r as.character(my_site)`.  The cost of the work is estimated at \$`r as.character(my_cost_estimate() %>% pull(cost_est_1000s) * 1000)` for a cost benefit of \$`r as.character(my_cost_estimate() %>% pull(cost_net) * 1000)`/linear m and \$`r as.character(my_cost_estimate() %>% pull(cost_area_net) * 1000)`/m^2^.

<br>

## Conclusion {-}

There is an estimated `r my_priority_info() %>% pull(upstream_habitat_length_km)`km of mainstem habitat upstream of crossing `r as.character(my_site)` with habitat in the areas surveyed upstream of the crossing rated as `r my_priority_info(loc = 'us') %>% pull(hab_value)` value for fry and juvenile salmonid rearing. Density of westslope cutthrout trout parr was higher in the site sampled downstream of the crossing when compared to the upstream site with fry, juvenile and adult fish captured downstren only. The `r my_overview_info() %>% pull(road_name)` is under tenure of the Ministry of Forests, Lands, Natural Resource Operations & Rural Development. The crossing was ranked as a `r my_priority_info() %>% pull(priority)` priority for proceeding to design for replacement with an open bottomed structure and consideration of remediation of PSCIS 50185 ([Appendix - Crossing 50185]) should be considered at the same time for maximum potential benefits.

\pagebreak

```{r tab-culvert-50181, eval = T}
print_tab_summary()

```

\pagebreak

```{r tab-habitat-summary-50181}
tab_hab_summary %>% 
  filter(Site  == my_site) %>% 
  select(-Site) %>% 
  my_kable(caption_text = paste0('Summary of habitat details for PSCIS crossing ', my_site, '.'))

```


<br>

```{r tab-fish-site-50181, eval=T}
my_caption <- paste0('Electrofishing sites for PSCIS crossing ', my_site, '.')

tab_fish_site() %>% 
  my_kable(caption_text = my_caption)
```

\pagebreak

```{r tab-fish-dens-50181, eval=T}
my_caption <- paste0('Westslope cuthrout trout densities (fish/100m2) for PSCIS crossing ', my_site, '.')

tab_fish_dens() %>% 
  my_kable(caption_text = my_caption)
```

<br>

```{r plot-fish-box-50181, fig.cap= my_caption, eval=T}
my_caption <- paste0('Fish densities (fish/100m2) for PSCIS crossing ', my_site, '.')

p <- plot_fish_box() 
p
```


<br>

```{r photo-50181-01, fig.cap= my_caption, out.width = photo_width, eval=T}
my_photo = pull_photo_by_str(str_to_pull = '_d1_')

my_caption = paste0('Typical habitat downstream of PSCIS crossing ', my_site, '.')

grid::grid.raster(get_img())

# knitr::include_graphics(get_img_path())
```

<br>

```{r photo-50181-02, fig.cap= my_caption, out.width = photo_width, eval=T}

# 
my_photo = pull_photo_by_str(str_to_pull = '_u1_')

my_caption = paste0('Typical habitat upstream of PSCIS crossing ', my_site, '.')

grid::grid.raster(get_img())
# knitr::include_graphics(get_img_path())
```

<br>

```{r photo-50181-03, fig.cap= my_caption, out.width = photo_width, eval=T}

# 
my_photo = pull_photo_by_str(str_to_pull = '_ufish')

my_caption = paste0('Westslope cutthrout trout captured upstream of PSCIS crossing ', my_site, '.')

grid::grid.raster(get_img())
# knitr::include_graphics(get_img_path())
```



