
```{r setup-197533, eval = F}
knitr::opts_chunk$set(echo=FALSE, message=FALSE, warning=FALSE)
source('R/packages.R')
source('R/functions.R')
source('R/functions-phase2.R')
source('R/tables-phase2.R')
source('R/tables.R')
source('R/extract-fish.R')
source('R/functions-fish.R')


```



```{r  load-data-197533}
my_site <- 197533
my_site2 <- 197559
my_site_xref <- 4600183
my_site2_xref <- 4600070
```


`r paste0('# Appendix - Crossings ', my_site, ' and ', my_site2)`{-} 
`r paste0('**', my_overview_info() %>% pull(road_name), ' and ', my_overview_info(site = my_site2) %>% pull(road_name), ' - ', my_overview_info() %>% pull(stream_name), '**')`


## Site Location {-}

PSCIS crossings `r as.character(my_site)` and `r as.character(my_site2)` are located north of Sparwood, BC on `r my_pscis_info() %>% pull(stream_name)`, approximately 600m and 725m upstream from the Elk River.  During 2020 surveys, two fords (PSCIS 197535 and 197536) were documented 700m and 2km upstream of crossing `r as.character(my_site2)` respectively.  Although several un-assessed crossings are modelled upstream of PSCIS 197536 (ford), review of aerial imagery indicates that the sole crossing upstream on the mainstem of Brule Creek is a bridge (modelled crossing 24706664) and the remaining upstream crossings are on small and/or very steep tributaries and unlikely to be blocking access to significant amounts of important habitat. Both Busato Road and Highway 43 are the responsibility of the Ministry of Transportation and Infrastructure.   


<br>

## Background {-}


At the crossing locations, Brule Creek is a 5th order stream with a watershed area upstream of the highway of approximately 87km^2^. The elevation of the watershed ranges from a maximum of 2800 to 1200m at  PSCIS crossing `r as.character(my_site)`. Brule Creek is known to contain westslope cutthrout trout, rainbow trout and bull trout downstream of the subject culverts and westslope cutthrout trout and rainbow trout above [@data_fish_obs]. On the south side of the upper watershed, at an elevation of 2000m, is the 5ha Josephine Lake (also known as Big Lake). The lake was stocked with westslope cutthrout trout from 1983 - 2000 [@data_fish_obs; @FishInventoriesData2020].

<br>

PSCIS stream crossings `r as.character(my_site)` and `r as.character(my_site2)` were ranked as high priorities for follow up with fish passage assessments and habitat confirmations due to the large size of the stream network upstream (`r my_bcfishpass() %>% pull(wct_network_km)`km) of the highway and because Brule Creek is a 5th order stream. The habitat confirmation was completed on `r my_pscis_info() %>% pull(date) %>% format(., format = "%B %d, %Y")`. A map of the watershed including areas surveyed is provided in Attachment 1 – Map [`r my_bcfishpass() %>% pull(map_tile_display_name)`](`r my_mapsheet()`).

<br>

## Stream Characteristics at Crossing {-}

At the time of the survey, both culverts were un-embedded, non-backwatered and considered barriers to upstream fish passage. PSCIS crossing `r as.character(my_site)` was `r my_pscis_info() %>% pull(diameter_or_span_meters)`m in diameter with a pipe length of `r my_pscis_info() %>% pull(length_or_width_meters)`m, a culvert slope of `r my_pscis_info() %>% pull(culvert_slope_percent)`%, a stream width ratio of `r my_pscis_info() %>% pull(stream_width_ratio)` and an outlet drop of `r my_pscis_info() %>% pull(outlet_drop_meters)`m (Table \@ref(tab:tab-culvert-197533)). PSCIS crossing `r as.character(my_site2)` was `r my_pscis_info(site = my_site2) %>% pull(diameter_or_span_meters)`m in diameter with a pipe length of `r my_pscis_info(site = my_site2) %>% pull(length_or_width_meters)`m, a culvert slope of `r my_pscis_info(site = my_site2) %>% pull(culvert_slope_percent)`%, a stream width ratio of `r my_pscis_info(site = my_site2) %>% pull(stream_width_ratio)` and an outlet drop of `r my_pscis_info(site = my_site2) %>% pull(outlet_drop_meters)`m (Table \@ref(tab:tab-culvert-197559)). Water temperature was `r my_habitat_info() %>% filter(rowname == 'temperature c') %>% pull(us)`$^\circ$C, pH was `r my_habitat_info() %>% filter(rowname == 'p h') %>% pull(us)` and conductivity was `r my_habitat_info() %>% filter(rowname == 'conductivity m s cm') %>% pull(us)`uS/cm.  

<br>


```{r eval=F}
##this is useful to get some comments for the report
hab_site %>% filter(site == my_site & location == 'us') %>% pull(comments)
hab_site %>% filter(site == my_site2 & location == 'us') %>% pull(comments)
my_priority <-  my_priority_info()

```

## Stream Characteristics Downstream of `r as.character(my_site)` {-}

The stream was surveyed downstream from the culvert for `r my_priority_info(loc = 'ds') %>% pull(survey_length_m)`m. Overall, total cover amount was rated as `r my_habitat_info2(loc = 'ds') %>% filter(rowname == 'total cover') %>% pull(v)` with  `r my_habitat_info2(loc = 'ds') %>% filter(v == 'dominant') %>% pull(rowname)` dominant. Cover was also present as `r my_habitat_info2(loc = 'ds') %>% filter(v == 'sub-dominant') %>% pull(rowname) %>% knitr::combine_words()` (Table \@ref(tab:tab-habitat-summary-197533), Figure \@ref(fig:photo-197533-01)). The average channel width was `r my_habitat_info3(loc = 'ds', row = 'avg channel width m')`m, the average wetted width was `r my_habitat_info3(loc = 'ds', row = 'avg wetted width m')`m and the average gradient was `r my_habitat_info3(loc = 'ds', row = 'average gradient percent')`%. The dominant substrate was `r my_habitat_info3(loc = 'ds', row = "bed material dominant")` with `r my_habitat_info3(loc = 'ds', row = "bed material subdominant")` subdominant. In the area surveyed, the stream channel appeared to be anthropogenically channelized and straightened with influence of adjacent livestock grazing/access negatively impacting stream banks due to loss of riparian vegetation and erosion. Habitat was rated as `r my_priority_info(loc = 'ds') %>%  pull(hab_value)` as it was considered an important migration corridor with moderate value habitat for fry/juvenile salmonid rearing.  

<br>

## Stream Characteristics Upstream of `r as.character(my_site)` and Downstream of `r as.character(my_site2)`{-}

The stream was surveyed upstream from `r as.character(my_site)` for `r my_priority_info(loc = 'us') %>% pull(survey_length_m)`m to `r as.character(my_site2)`. Overall, total cover amount was rated as `r my_habitat_info2(loc = 'us') %>% filter(rowname == 'total cover') %>% pull(v)` with  `r my_habitat_info2(loc = 'us') %>% filter(v == 'dominant') %>% pull(rowname)` dominant. Cover was also present as `r my_habitat_info2(loc = 'us') %>% filter(v == 'sub-dominant') %>% pull(rowname) %>% knitr::combine_words()` (Table \@ref(tab:tab-habitat-summary-197533), Figure \@ref(fig:photo-197533-02)). The average channel width was `r my_habitat_info3(loc = 'us', row = 'avg channel width m')`m, the average wetted width was `r my_habitat_info3(loc = 'us', row = 'avg wetted width m')`m and the average gradient was `r my_habitat_info3(loc = 'us', row = 'average gradient percent')`%.  There was a large pool at the outlet of crossing `r as.character(my_site2)` containing approximately nine westslope cutthrout trout with lengths of five of the fish estimated at 200mm and four at 300mm (Figure \@ref(fig:photo-197533-03)).  The outlet pool was depth was >1m with gravels suitable for spawning at the tailout.  Habitat value was rated as `r my_priority_info(loc = 'us') %>% pull(hab_value)` for resident and fluvial salmonid rearing and spawning. 


<br>

## Stream Characteristics Upstream of `r as.character(my_site2)`{-}

The stream was surveyed immediately upstream from `r as.character(my_site2)` for 1200m and then another 600m beginning at the ford (PSCIS 197536) 2km upstream for a total of `r my_priority_info(sit = my_site2, loc = 'us') %>% pull(survey_length_m)`m. The channel was dewatered immediately upstream of Highway 43 with intermittent pools only to a distance approximately 670 m upstream.  Upstream of the dewatered area, stream flows increased with increasing upstream distance.  Overall, total cover amount was rated as `r my_habitat_info2(sit = my_site2, loc = 'us') %>% filter(rowname == 'total cover') %>% pull(v)` with  `r my_habitat_info2(sit = my_site2, loc = 'us') %>% filter(v == 'dominant') %>% pull(rowname)` dominant. Cover was also present as `r my_habitat_info2(sit = my_site2,, loc = 'us') %>% filter(v == 'sub-dominant') %>% pull(rowname) %>% knitr::combine_words()` (Table \@ref(tab:tab-habitat-summary-197533), Figure \@ref(fig:photo-197533-02)). The average channel width was `r my_habitat_info3(loc = 'us', row = 'avg channel width m')`m, the average wetted width was `r my_habitat_info3(sit = my_site2, loc = 'us', row = 'avg wetted width m')`m and the average gradient was `r my_habitat_info3(sit = my_site2, loc = 'us', row = 'average gradient percent')`%.  Overall, habitat upstream of `r as.character(my_site2)` was rated as `r my_priority_info(sit = my_site2, loc = 'us') %>% pull(hab_value)` value habitat present suitable for resident and fluvial salmonid rearing and spawning. 

<br>

## Fish Sampling {-}

To assess potential impacts of the culvert on fisheries values in the stream, electrofishing was conducted upstream of the crossing.  Sampling was not conducted downstream of crossing `r as.character(my_site2)` as bull trout presence was suspected and it was determined that spawners and/or eggs may have been present.  One site was sampled upstream with no fish captured (Figure \@ref(fig:photo-032)).  Raw results are included in digital format as [Attachment 2](https://github.com/NewGraphEnvironment/fish_passage_elk_2020_reporting/raw/master/data/habitat_confirmations.xls) and summarized in Table \@ref(tab:tab-fish-site-197533).


<br>

## Structure Remediation and Cost Estimate {-}

As properties on Busato Road could be accessed from either side of the road, removal of PSCIS `r as.character(my_site)` could be explored as an option for providing access to the 125m of habitat located upstream and below PSCIS `r as.character(my_site)`.  However, to facilitate an estimate of "worst case scenario", costs for replacement of both crossing `r as.character(my_site)` and `r as.character(my_site)` with bridges are estimated at \$`r as.character(my_cost_estimate() %>% pull(cost_est_1000s) * 1000)` and \$`r as.character(my_cost_estimate(site = my_site2) %>% pull(cost_est_1000s) * 1000)` respectively.  The combined cost benefit of replacements are estimated at `r paste0('$', as.character(my_cost_estimate() %>% pull(cost_net) * 1000 + my_cost_estimate(site = my_site2) %>% pull(cost_net) * 1000))`/linear m and \$`r as.character(my_cost_estimate() %>% pull(cost_area_net) * 1000  + my_cost_estimate(site = my_site2) %>% pull(cost_area_net) * 1000)`/m^2^.


<br>



## Conclusion {-}

There is `r my_priority_info() %>% pull(upstream_habitat_length_km)`km of habitat upstream of crossing `r as.character(my_site2)` and another `r my_priority_info(sit = my_site2) %>% pull(upstream_habitat_length_km)`km upstream of `r as.character(my_site2)` rated as `r my_priority_info(sit = my_site, loc = 'us') %>% pull(hab_value)` value for resident and fluvial salmonid rearing/spawning. The lack of fish captured upstream of the crossing is not an indication of a lack of fish presence but may be indicative of low population densities and/or restricted access due to  downstream culverts and dewatering. Although an interim ranking for remediation at the crossings was assessed as `r my_priority_info() %>% pull(priority)` priority to proceeding to designs for both crossings, follow up to determine the extent of dewatering during higher flow periods is recommended.  Although unconfirmed at the time of reporting, the 670m section of stream located immediately upstream of Highway 43 that was flowing subsurface at the time of the survey, very likely flows above ground during high and peak flows when adult westslope cutthrout trout display a general pattern of upstream movement to spawning areas [@schweigert_etal2017COSEWICassessment].

<br>

```{r tab-culvert-197533, eval = T}
print_tab_summary(site_photo_id = my_site_xref)

```

<br>

```{r tab-culvert-197559, eval = T}
##had to pull this object out of the table push for some reason
print_tab_summary(site = my_site2, site_photo_id = my_site2_xref)

```

<br>

```{r tab-habitat-summary-197533}
tab_hab_summary %>% 
  filter(Site  == my_site | Site  == my_site2) %>% 
  # select(-Site) %>% 
  my_kable(caption_text = paste0('Summary of habitat details for PSCIS crossings ', my_site, ' and ', my_site2, '.'))

```

<br>

```{r tab-fish-site-197533, eval=T}
my_caption <- paste0('Electrofishing site upstream of PSCIS crossing ', my_site2, '.')

tab_fish_site(sit = my_site2) %>% 
  my_kable(caption_text = my_caption)
```


<br>

```{r photo-197533-01, fig.cap= my_caption, eval=T}
my_photo = '20200918_111231_k_d_01_.JPG'

my_caption = paste0('Typical habitat downstream of PSCIS crossing ', my_site, '.')

grid::grid.raster(get_img(site = my_site_xref))

# knitr::include_graphics(get_img_path())
```

<br>

```{r photo-197533-02, fig.cap= my_caption, eval=T}
my_photo = '20200918_114535_k_u01_.JPG'

my_caption = paste0('Typical habitat upstream of PSCIS crossing ', my_site, ' and downstream of PSCIS crossing ', my_site, '.')

grid::grid.raster(get_img(site = my_site_xref))
# knitr::include_graphics(get_img_path())
```

<br>

```{r photo-197533-03, fig.cap= my_caption, eval=T}
my_photo = 'IMG_3074_k_d01_wct.JPG'

my_caption = paste0('Westlope cutthrout trout in outlet pool downstream of PSCIS crossing ', my_site2, '.')

grid::grid.raster(get_img(site = my_site2_xref))
# knitr::include_graphics(get_img_path())
```

<br>

```{r photo-197533-04, fig.cap= my_caption, eval=T}
my_photo = 'IMG_3080_k_u01_.JPG'

my_caption = paste0('Subsurface flow area upstream of PSCIS crossing ', my_site2, '.')

grid::grid.raster(get_img(site = my_site2_xref))
# knitr::include_graphics(get_img_path())
```

<br>

```{r photo-197533-05, fig.cap= my_caption, eval=T}
my_photo = 'IMG_3102_K_u02_.JPG'

my_caption = paste0('Typical habitat above dewatered area upstream of PSCIS crossing ', my_site, '.')

grid::grid.raster(get_img(site = my_site2_xref))
# knitr::include_graphics(get_img_path())
```

<br>

```{r photo-197533-06, fig.cap= my_caption, eval=T}
my_photo = 'IMG_3110_k_u03_ef_.JPG'

my_caption = paste0('Habitat within electrofishing site upstream of PSCIS crossing ', my_site, '.')

grid::grid.raster(get_img(site = my_site2_xref))
# knitr::include_graphics(get_img_path())
```

