
```{r setup-02, eval = F}
knitr::opts_chunk$set(echo=FALSE, message=FALSE, warning=FALSE)
source('R/packages.R')
source('R/functions.R')
source('R/functions-phase2.R')
source('R/tables-phase2.R')
source('R/tables.R')
source('R/extract-fish.R')
source('R/functions-fish.R')


```



```{r  load-data-02}
my_site <- 50159
```


`r paste0('# Appendix - Crossing ', my_site)`{-} 
`r paste0('**', my_overview_info() %>% pull(road_name), ' - ', my_overview_info() %>% pull(stream_name), '**')`


## Site Location {-}

Crossing `r as.character(my_site)` is located on a `r my_pscis_info() %>% pull(stream_name)`, approximately `r my_priority_info(loc = 'ds') %>% pull(survey_length_m)`m upstream from the confluence with Lizard Creek. This crossing has also been recorded in PSCIS as crossing 62510.  The culvert is located at approximately 5.5km up Island Lake Lodge Road which is an extension of Mt.Fernie Park Road.  The area is accessed off of Highway 3 within Fernie city limits and is a popular recreational destination for hikers, skiers and mountain bikers.  Island Lake Lodge is located at 1400m of elevation near Island Lake and is a year round tourist destination providing accommodations, guided hiking and backcountry catskiing for clients. The subject stream is not mapped in the freshwater atlas stream layer and may have been diverted as part of a micro-hydro facilty for Island Lake Lodge.  A small building that may be a generating station was observed on aerial imagery approximately 1500m upstream of the road.


<br>

## Background {-}

At the crossing location, the stream had good flow and is located within an area of old growth cedar adjacent to a recreation trail. At the time of the survey the stream was the highest volume tributary to Lizard Creek located on the east side of the valley. No fisheries information was available for the stream [@moeStreamInventorySample], however, Lizard Creek supports westslope cutthroat trout, bull trout, mountain whitefish, brook trout, longnose sucker and longnose dace [@data_fish_obs]. @elkriveralliance2020ElkRiver conducted redd surveys in Lizard Creek in 2019 along with Morrissey Creek, Coal Creek and Forsyth Creek.  A total of 55 redds were observed within a 2.4km section of Lizard Creek comprising the largest densities of redds of the four tributaries surveyed (22.9 redds/km).

<br>

PSCIS stream crossing `r as.character(my_site)` was ranked as a high priority for follow up with habitat confirmation due to the relatively large size of the stream recorded in PSCIS (channel width = 3.5m) relative to other tributary streams in the Lizard Creek watershed and because it was rated as containing high value habitat by @vastFishPassage2013 and @grainger2011FishPassage2011. The habitat confirmation was completed on `r my_pscis_info() %>% pull(date) %>% format(., format = "%B %d, %Y")`. A map of the watershed including areas surveyed is provided in Attachment 1 – Map [`r my_bcfishpass() %>% pull(map_tile_display_name)`](`r my_mapsheet()`).

<br>

## Stream Characteristics at Crossing {-}

At the time of the survey, the un-embedded and non-backwatered `r my_pscis_info() %>% pull(diameter_or_span_meters)`m diameter crossing was considered a barrier to upstream fish passage with a pipe length of `r my_pscis_info() %>% pull(length_or_width_meters)`m, a culvert slope of `r my_pscis_info() %>% pull(culvert_slope_percent)`%, a stream width ratio of `r my_pscis_info() %>% pull(stream_width_ratio)` and an outlet drop of `r my_pscis_info() %>% pull(outlet_drop_meters)`m (Table \@ref(tab:tab-culvert-020)). Water temperature was `r my_habitat_info() %>% filter(rowname == 'temperature c') %>% pull(us)`$^\circ$C, pH was `r my_habitat_info() %>% filter(rowname == 'p h') %>% pull(us)` and conductivity was `r my_habitat_info() %>% filter(rowname == 'conductivity m s cm') %>% pull(us)`uS/cm.  

<br>


```{r eval=F}
##this is useful to get some comments for the report
hab_site %>% filter(site == my_site & location == 'us') %>% pull(comments)

```
## Stream Characteristics Downstream {-}

The stream was surveyed downstream from the culvert for `r my_priority_info(loc = 'ds') %>% pull(survey_length_m)`m to Lizard Creek. Overall, total cover amount was rated as `r my_habitat_info2(loc = 'ds') %>% filter(rowname == 'total cover') %>% pull(v)` with  `r my_habitat_info2(loc = 'ds') %>% filter(v == 'dominant') %>% pull(rowname)` dominant. Cover was also present as `r my_habitat_info2(loc = 'ds') %>% filter(v == 'sub-dominant') %>% pull(rowname) %>% knitr::combine_words()` (Table \@ref(tab:tab-habitat-summary-010), Figure \@ref(fig:photo-001)). The average channel width was `r my_habitat_info3(loc = 'ds', row = 'avg channel width m')`m, the average wetted width was `r my_habitat_info3(loc = 'ds', row = 'avg wetted width m')`m and the average gradient was `r my_habitat_info3(loc = 'ds', row = 'average gradient percent')`%.  The dominant substrate was `r my_habitat_info3(loc = 'ds', row = "bed material dominant")` with `r my_habitat_info3(loc = 'ds', row = "bed material subdominant")` subdominant. Some small pools and steps of 0.2 - 0.6m in height were present throughout the area surveyed. Large woody debris steps ranging from 0.4 - 0.8m high were spaced sporadically throughout area surveyed.  The stream contained relatively flatter sections from 3 - 8% and steeper sections of 15 - 18% for first 350m above culvert (UTM: 11 U 632810 5484842) where the stream becomes too steep for upstream salmonid passage (35%). Habitat value was rated as moderate for salmonid fry/juvenile rearing and high value habitat for spawning.  

<br>

## Stream Characteristics Upstream {-}

The stream was surveyed upstream from the culvert for `r my_priority_info(loc = 'us') %>% pull(survey_length_m)`m. Overall, total cover amount was rated as `r my_habitat_info2(loc = 'us') %>% filter(rowname == 'total cover') %>% pull(v)` with  `r my_habitat_info2(loc = 'us') %>% filter(v == 'dominant') %>% pull(rowname)` dominant. Cover was also present as `r my_habitat_info2(loc = 'us') %>% filter(v == 'sub-dominant') %>% pull(rowname) %>% knitr::combine_words()` (Table \@ref(tab:tab-habitat-summary-010), Figure \@ref(fig:photo-002)). The average channel width was `r my_habitat_info3(loc = 'us', row = 'avg channel width m')`m, the average wetted width was `r my_habitat_info3(loc = 'us', row = 'avg wetted width m')`m and the average gradient was `r my_habitat_info3(loc = 'us', row = 'average gradient percent')`%.  There were abundant gravels suitable for resident and fluvial westslope cutthrout trout spawning.  Frequent pools to 40cm deep were present and associated with small and large woody debris. Habitat was rated as `r my_priority_info(loc = 'us') %>% pull(hab_value)` value for salmonid rearing and spawning. 


<br><br>

## Structure Remediation and Cost Estimate {-}

Structure replacement with a bridge is recommended to provide access to the habitat located upstream of PSCIS crossing `r as.character(my_site)`.  The cost for the work is estimated at \$`r as.character(my_cost_estimate() %>% pull(cost_est_1000s) * 1000)` for a cost benefit of \$`r as.character(my_cost_estimate() %>% pull(cost_net) * 1000)`/linear m and \$`r as.character(my_cost_estimate() %>% pull(cost_area_net) * 1000)`/m^2^.

<br><br>


## Conclusion {-}

There is `r my_priority_info() %>% pull(upstream_habitat_length_km)`km of habitat upstream of crossing `r as.character(my_site)` within old growth cedar riparian forest before the stream becomes too steep for upstream fish migration (>20%).  The habitat upstream of the crossing was rated as high value for salmonid rearing and spawning with good flows, large woody debris throughout and abundant gravels suitable for spawning. The subject stream is not mapped in the freshwater atlas stream layer and may have been diverted from upstream flows as part of a micro-hydro facilty for Island Lake Lodge.  The road may be permitted to Island Lake Lodge or solely the responsibility of the Ministry of Forests, Lands, Natural Resource Operations & Rural Development. The crossing was ranked as a `r my_priority_info() %>% pull(priority)` priority for proceeding to design for replacement with an open bottomed structure.  Elecrofishing upstream and downstream of the crossing is recommended to determine utilization of habitat before and after restorative measures.


\pagebreak

```{r tab-culvert-020, eval = T}
##had to pull this object out of the table push for some reason
print_tab_summary()

```

\pagebreak

```{r tab-habitat-summary-020}
tab_hab_summary %>% 
  filter(Site  == my_site) %>% 
  select(-Site) %>% 
  my_kable(caption_text = paste0('Summary of habitat details for PSCIS crossing ', my_site, '.'))

```


<br>

```{r photo-020, fig.cap= my_caption, out.width = photo_width, eval=T}
my_photo = '20200922_122212_k_.JPG'

my_caption = paste0('Typical habitat downstream of PSCIS crossing ', my_site, '.')

grid::grid.raster(get_img())

# knitr::include_graphics(get_img_path())
```

<br>

```{r photo-021, fig.cap= my_caption, out.width = photo_width, eval=T}

# 
my_photo = 'IMG_3287_k_.JPG'

my_caption = paste0('Typical habitat upstream of PSCIS crossing ', my_site, '.')

grid::grid.raster(get_img())
# knitr::include_graphics(get_img_path())
```

