---
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup-50181, eval = F}
knitr::opts_chunk$set(echo=FALSE, message=FALSE, warning=FALSE)
source('R/packages.R')
source('R/functions.R')
source('R/functions-phase2.R')
source('R/tables-phase2.R')
source('R/tables.R')
source('R/extract-fish.R')
source('R/functions-fish.R')


```



```{r  load-data-50181}
my_site <- 50181

```


```{r eval = F}
my_bcfish <- my_bcfishpass() ##can be useful to look at
```


`r paste0('# Appendix - Crossing ', my_site)`{-} 
`r paste0('**', my_overview_info() %>% pull(road_name), ' - ', my_overview_info() %>% pull(stream_name), '**')`


## Site Location {-}

Crossing `r as.character(my_site)` is located on a `r my_pscis_info() %>% pull(stream_name)`, approximately `r my_priority_info(loc = 'ds') %>% pull(survey_length_m)`m upstream from the confluence with Morrisey Creek. The crossing is located on River Road just south of Morrissey approximately 15km south of Fernie.  The road is accessed off of Lodgepole FSR via the Morrissey Bridge over the Elk River adjacent to Highway 3. River Road is a gravel forest tenure road (forest file id `r my_bcfishpass() %>%  pull(forest_file_id)` with active log hauling at the time of the survey. 


<br>

## Background {-}

At the crossing location, the stream is 4th order with a watershed area upstream of the road of approximately 12km^2^. The elevation of the watershed ranges from a maximum of 2000 to 970m at the culvert. Upstream of the crossing, there are no anthropogenic barriers on the mainstem however PSCIS crossing 50181 (also recorded as PSCIS 103033) is documented as a barrier located on a significantly sized tributary entering the stream from the north-east approximately 2.8km upstream of River Road. A wetland type area is mapped at the top of the watershed (area `r my_bcfishpass() %>%  pull(wetland_area_ha)`ha). No fisheries information was available for the stream [@moeStreamInventorySample] however westslope cutthroat trout, bull trout, mountain whitefish, brook trout, longnose sucker and longnose dace have been recorded downstream in Morrissey Creek [@data_fish_obs].   

<br>

PSCIS stream crossing `r as.character(my_site)` was ranked as a high priority for follow up with habitat confirmation due to the large size of the stream relative to other tributary streams in the watershed, the previously rated moderate value habitat as rated by @vastFishPassage2013. The habitat confirmation was completed on `r my_pscis_info() %>% pull(date) %>% format(., format = "%B %d, %Y")`. A map of the watershed including areas surveyed is provided in Attachment 1 – Map [`r my_bcfishpass() %>% pull(map_tile_display_name)`](`r my_mapsheet()`).

<br>

## Stream Characteristics at Crossing {-}

At the time of the survey, the un-embedded and non-backwatered `r my_pscis_info() %>% pull(diameter_or_span_meters)`m diameter crossing was considered a barrier to upstream fish passage with a pipe length of `r my_pscis_info() %>% pull(length_or_width_meters)`m, a culvert slope of `r my_pscis_info() %>% pull(culvert_slope_percent)`%, a stream width ratio of `r my_pscis_info() %>% pull(stream_width_ratio)` and an outlet drop of `r my_pscis_info() %>% pull(outlet_drop_meters)`m (Table \@ref(tab:tab-culvert-50181)). Water temperature was `r my_habitat_info() %>% filter(rowname == 'temperature c') %>% pull(us)`$^\circ$C, pH was `r my_habitat_info() %>% filter(rowname == 'p h') %>% pull(us)` and conductivity was `r my_habitat_info() %>% filter(rowname == 'conductivity m s cm') %>% pull(us)`uS/cm. It appeared as though fish passage restoration works had taken place at the site historically as there were what appeared to be boulder riffle structures installed downstream of the crossing.  The structures appeared to be effective at reducing the outlet drop size but had created a rock drop (0.4m) and were not resulting in backwatering of the culvert. 

<br>


```{r eval=F}
##this is useful to get some comments for the report
hab_site %>% filter(site == my_site & location == 'ds') %>% pull(comments)
my_hab_info <- my_habitat_info()
```
## Stream Characteristics Downstream {-}

The stream was surveyed downstream from the culvert for `r my_priority_info(loc = 'ds') %>% pull(survey_length_m)`m to the confluence with Morrisey Creek. The mouth of the stream is located approximately 500m upstream from the confluence of Morrissey Creek and the Elk River. Overall, total cover amount was rated as `r my_habitat_info2(loc = 'ds') %>% filter(rowname == 'total cover') %>% pull(v)` with  `r my_habitat_info2(loc = 'ds') %>% filter(v == 'dominant') %>% pull(rowname)` dominant. Cover was also present as `r my_habitat_info2(loc = 'ds') %>% filter(v == 'sub-dominant') %>% pull(rowname) %>% knitr::combine_words()` (Table \@ref(tab:tab-habitat-summary-50181), Figure \@ref(fig:photo-50181-01)). The average channel width was `r my_habitat_info3(loc = 'ds', row = 'avg channel width m')`m, the average wetted width was `r my_habitat_info3(loc = 'ds', row = 'avg wetted width m')`m and the average gradient was `r my_habitat_info3(loc = 'ds', row = 'average gradient percent')`%.  The dominant substrate was `r my_habitat_info3(loc = 'ds', row = "bed material dominant")` with `r my_habitat_info3(loc = 'ds', row = "bed material subdominant")` subdominant.  There were frequent pools formed by small and large woody debris ranging from 0.3 - 0.75m in depth (average residual depth = `r my_habitat_info3(loc = 'ds', row = "average residual pool depth m")`m).  Pockets of small gravels suitable for resident salmonid spawning were also present. Habitat value was rated as `r my_priority_info(loc = 'ds') %>%  pull(hab_value)` with good potential for fry/juvenile salmonid rearing.


```{r eval=F}
##this is useful to get some comments for the report
hab_site %>% filter(site == my_site & location == 'us') %>% pull(comments)
my_hab_info <- my_habitat_info()
```


<br>

## Stream Characteristics Upstream {-}

The stream was surveyed upstream from the culvert for `r my_priority_info(loc = 'us') %>% pull(survey_length_m)`m. Overall, total cover amount was rated as `r my_habitat_info2(loc = 'us') %>% filter(rowname == 'total cover') %>% pull(v)` with  `r my_habitat_info2(loc = 'us') %>% filter(v == 'dominant') %>% pull(rowname)` dominant. Cover was also present as `r my_habitat_info2(loc = 'us') %>% filter(v == 'sub-dominant') %>% pull(rowname) %>% knitr::combine_words()` (Table \@ref(tab:tab-habitat-summary-50181), Figure \@ref(fig:photo-50181-02)). The average channel width was `r my_habitat_info3(loc = 'us', row = 'avg channel width m')`m, the average wetted width was `r my_habitat_info3(loc = 'us', row = 'avg wetted width m')`m and the average gradient was `r my_habitat_info3(loc = 'us', row = 'average gradient percent')`%.  The dominant substrate was `r my_habitat_info3(loc = 'us', row = "bed material dominant")` with `r my_habitat_info3(loc = 'us', row = "bed material subdominant")` subdominant. The stream had good flows with fry observed throughout the area surveyed.  Pools to 0.6m deep (average residual depth =`r my_habitat_info3(loc = 'us', row = "average residual pool depth m")`m) were present with pockets of gravel suitable for spawning throughout.  Infrequent large woody debris jams to 0.5m high were also observed. There were frequent areas of gravels suitable for resident westslope cutthrout trout spawning. Habitat value was rated as `r my_priority_info(loc = 'us') %>% pull(hab_value)` for fry and juvenile westslope cutthrout trout rearing. 

<br>

## Fish Sampling {-}

To assess potential impacts of the culvert on fish densities in the stream electrofishing was conducted upstream and downstream of the crossing.  Five sites were sampled downstream and five sites  were sampled upstream. A total of `r tab_fish_summary %>% filter(site_id == paste0(my_site, '_us') & species_code == 'WCT') %>% pull(count_fish)` westslope cutthrout trout and `r tab_fish_summary %>% filter(site_id == paste0(my_site, '_us') & species_code == 'EB') %>% pull(count_fish)` eastern brook trout were captured upstream with `r tab_fish_summary %>% filter(site_id == paste0(my_site, '_ds') & species_code == 'WCT') %>% pull(count_fish)` westslope cutthrout trout and `r tab_fish_summary %>% filter(site_id == paste0(my_site, '_ds') & species_code == 'EB') %>% pull(count_fish)` eastern brook trout captured downstream (Figure \@ref(fig:photo-50181-02)).  Raw results are included in digital format as [Attachment 2](https://github.com/NewGraphEnvironment/fish_passage_elk_2020_reporting/raw/master/data/habitat_confirmations.xls) and summarized in Tables \@ref(tab:tab-fish-site-50181) - \@ref(tab:tab-fish-dens-50181) and Figure \@ref(fig:plot-fish-box-50181).

<br>

## Structure Remediation and Cost Estimate {-}

Structure replacement with an open bottomed structure is recommended to provide access to the habitat located upstream of PSCIS crossing `r as.character(my_site)`.  The cost of the work is estimated at \$`r as.character(my_cost_estimate() %>% pull(cost_est_1000s) * 1000)` for a cost benefit of \$`r as.character(my_cost_estimate() %>% pull(cost_net) * 1000)`/linear m and \$`r as.character(my_cost_estimate() %>% pull(cost_area_net) * 1000)`/m^2^.

<br>

## Conclusion {-}

There is an estimated `r my_priority_info() %>% pull(upstream_habitat_length_km)`km of mainstem habitat upstream of crossing `r as.character(my_site)` with habitat in the areas surveyed upstream of the crossing rated as high value for fry and juvenile salmonid rearing. Fish sampling results indicated that densities of westslope cutthrout trout fry and parr were similiar upstream and downstream of the crossing with a general trend of higher densities below the crossing for both life stages. River Road is under tenure of the Ministry of Forests, Lands, Natural Resource Operations & Rural Development. The crossing was ranked as a `r my_priority_info() %>% pull(priority)` priority for proceeding to design for replacement with an open bottomed structure.

\pagebreak

```{r tab-culvert-50181, eval = T}
print_tab_summary()

```

\pagebreak

```{r tab-habitat-summary-50181}
tab_hab_summary %>% 
  filter(Site  == my_site) %>% 
  select(-Site) %>% 
  my_kable(caption_text = paste0('Summary of habitat details for PSCIS crossing ', my_site, '.'))

```


<br>

```{r tab-fish-site-50181, eval=T}
my_caption <- paste0('Electrofishing sites for PSCIS crossing ', my_site, '.')

tab_fish_site() %>% 
  my_kable(caption_text = my_caption)
```

\pagebreak

```{r tab-fish-dens-50181, eval=T}
my_caption <- paste0('Westslope cuthrout trout densities (fish/100m2) for PSCIS crossing ', my_site, '.')

tab_fish_dens() %>% 
  my_kable(caption_text = my_caption)
```

<br>

```{r plot-fish-box-50181, fig.cap= my_caption, eval=T}
my_caption <- paste0('Fish densities (fish/100m2) for PSCIS crossing ', my_site, '.')

p <- plot_fish_box() 
p
```


<br>

```{r photo-50181-01, fig.cap= my_caption, out.width = photo_width, eval=F}
my_photo = '20200921_114215_k_.JPG'

my_caption = paste0('Typical habitat downstream of PSCIS crossing ', my_site, '.')

grid::grid.raster(get_img())

# knitr::include_graphics(get_img_path())
```

<br>

```{r photo-50181-02, fig.cap= my_caption, out.width = photo_width, eval=F}

# 
my_photo = 'IMG_3248_k_.JPG'

my_caption = paste0('Typical habitat upstream of PSCIS crossing ', my_site, '.')

grid::grid.raster(get_img())
# knitr::include_graphics(get_img_path())
```

<br>

```{r photo-50181-03, fig.cap= my_caption, out.width = photo_width, eval=F}

# 
my_photo = 'IMG_3254_k_.JPG'

my_caption = paste0('Westslope cutthrout trout captured upstream of PSCIS crossing ', my_site, '.')

grid::grid.raster(get_img())
# knitr::include_graphics(get_img_path())
```



