
```{r setup-09, eval = F}
knitr::opts_chunk$set(echo=FALSE, message=FALSE, warning=FALSE)
source('R/packages.R')
source('R/functions.R')
source('R/functions-phase2.R')
source('R/tables-phase2.R')
source('R/tables.R')
source('R/extract-fish.R')
source('R/functions-fish.R')


```



```{r load-data-09}
my_site <- 197555
my_site_xref <- 4605732

```


```{r eval = F}
my_bcfish <- my_bcfishpass() ##can be useful to look at
my_pscis <- my_pscis_info()
```


`r paste0('# Appendix - Crossing ', my_site)`{-} 
`r paste0('**', my_overview_info() %>% pull(road_name), ' - ', my_overview_info() %>% pull(stream_name), '**')`


## Site Location {-}

Crossing `r as.character(my_site)` is located on a `r my_pscis_info() %>% pull(stream_name)`, approximately 900m upstream from the confluence with the Elk River. The crossing is located on the Elk River FSR approximately 10km north of Elkford, BC. The Elk River FSR is a gravel forest tenure road (forest file id `r my_bcfishpass() %>%  pull(forest_file_id)`) with active log hauling at the time of the survey. 


<br>

## Background {-}

At the crossing location, the stream is 4th order with a watershed area upstream of the road of approximately 17km^2^. The elevation of the watershed ranges from a maximum of 2900 to 1400m at the culvert. Upstream of the crossing, there are no anthropogenic barriers on the mainstem.  Although there are two modelled crossings on small tributaries that enter the stream approximately 1.5km upstream of the road, machinery operators encountered onsite indicated that they had been removed during road deactivation.  No fisheries information was available for the stream [@moeStreamInventorySample] however westslope cutthroat trout, bull trout, mountain whitefish, brook trout, longnose sucker and longnose dace (among other species) have been recorded downstream in the Elk River [@data_fish_obs].   

<br>

PSCIS stream crossing `r as.character(my_site)` was ranked as a high priority for follow up with habitat confirmation due to the large amount of habitat modelled upstream of the crossing. The habitat confirmation was completed on `r my_pscis_info() %>% pull(date) %>% format(., format = "%B %d, %Y")`. A map of the watershed including areas surveyed is provided in Attachment 1 – Map [`r my_bcfishpass() %>% pull(map_tile_display_name)`](082J.103).

<br>

## Stream Characteristics at Crossing {-}

At the time of the survey, the un-embedded and non-backwatered `r my_pscis_info() %>% pull(diameter_or_span_meters)`m diameter crossing was considered a barrier to upstream fish passage with a pipe length of `r my_pscis_info() %>% pull(length_or_width_meters)`m, a culvert slope of `r my_pscis_info() %>% pull(culvert_slope_percent)`%, a stream width ratio of `r my_pscis_info() %>% pull(stream_width_ratio)` and an outlet drop of `r my_pscis_info() %>% pull(outlet_drop_meters)`m (Table \@ref(tab:tab-culvert-090)). Water temperature was `r my_habitat_info() %>% filter(rowname == 'temperature c') %>% pull(us)`$^\circ$C, pH was `r my_habitat_info() %>% filter(rowname == 'p h') %>% pull(us)` and conductivity was `r my_habitat_info() %>% filter(rowname == 'conductivity m s cm') %>% pull(us)`uS/cm.  

<br>


```{r eval=F}
##this is useful to get some comments for the report
hab_site %>% filter(site == my_site & location == 'ds') %>% pull(comments)
my_hab_info <- my_habitat_info()
my_pscis <- my_pscis_info() 
```
## Stream Characteristics Downstream {-}

The stream was surveyed downstream from the culvert for `r my_priority_info(loc = 'ds') %>% pull(survey_length_m)`m. Overall, total cover amount was rated as `r my_habitat_info2(loc = 'ds') %>% filter(rowname == 'total cover') %>% pull(v)` with  `r my_habitat_info2(loc = 'ds') %>% filter(v == 'dominant') %>% pull(rowname)` dominant. Cover was also present as `r my_habitat_info2(loc = 'ds') %>% filter(v == 'sub-dominant') %>% pull(rowname) %>% knitr::combine_words()` (Table \@ref(tab:tab-habitat-summary-090), Figure \@ref(fig:photo-090)). The average channel width was `r my_habitat_info3(loc = 'ds', row = 'avg channel width m')`m, the average wetted width was `r my_habitat_info3(loc = 'ds', row = 'avg wetted width m')`m and the average gradient was `r my_habitat_info3(loc = 'ds', row = 'average gradient percent')`%.  The dominant substrate was `r my_habitat_info3(loc = 'ds', row = "bed material dominant")` with `r my_habitat_info3(loc = 'ds', row = "bed material subdominant")` subdominant .  There were frequent pools formed by small and large woody debris ranging from 0.3 - 0.75m in depth (average residual depth = `r my_habitat_info3(loc = 'ds', row = "average residual pool depth m")`m). Frequent pockets of gravels suitably sized for resident and fluvial salmonid spawning were also present. Habitat value was rated as `r my_priority_info(loc = 'ds') %>%  pull(hab_value)` with good potential for fry/juvenile salmonid rearing.


```{r eval=F}
##this is useful to get some comments for the report
hab_site %>% filter(site == my_site & location == 'us') %>% pull(comments)
my_hab_info <- my_habitat_info()
```


<br>

## Stream Characteristics Upstream {-}

The stream was surveyed upstream from the culvert for `r my_priority_info(loc = 'us') %>% pull(survey_length_m)`m. Overall, total cover amount was rated as `r my_habitat_info2(loc = 'us') %>% filter(rowname == 'total cover') %>% pull(v)` with  `r my_habitat_info2(loc = 'us') %>% filter(v == 'dominant') %>% pull(rowname)` as the dominant type. Cover was also present as `r my_habitat_info2(loc = 'us') %>% filter(v == 'sub-dominant') %>% pull(rowname) %>% knitr::combine_words()` (Table \@ref(tab:tab-habitat-summary-090), Figure \@ref(fig:photo-091)). The average channel width was `r my_habitat_info3(loc = 'us', row = 'avg channel width m')`m, the average wetted width was `r my_habitat_info3(loc = 'us', row = 'avg wetted width m')`m and the average gradient was `r my_habitat_info3(loc = 'us', row = 'average gradient percent')`%.  The dominant substrate was `r my_habitat_info3(loc = 'us', row = "bed material dominant")` with `r my_habitat_info3(loc = 'us', row = "bed material subdominant")` subdominant. The stream had good flows, pools to 0.6m deep (average residual depth = `r my_habitat_info3(loc = 'us', row = "average residual pool depth m")`m) were present and pockets of gravel suitable for spawning were observed throughout.  Infrequent large woody debris jams to 0.5m high were also noted. Habitat value was rated as `r my_priority_info(loc = 'us') %>% pull(hab_value)` for fry and juvenile westslope cutthrout trout and bull trout rearing. 

<br>

## Fish Sampling {-}

To assess potential impacts of the culvert on fish communities in the stream, electrofishing was conducted upstream and downstream of the crossing.  One 150m site was sampled upstream and one 315m site was sampled downstream. One bull trout was captured downstream and no fish were captured upstream (Figure \@ref(fig:photo-092)).  Raw results are included in digital format as [Attachment 2](https://github.com/NewGraphEnvironment/fish_passage_elk_2020_reporting/raw/master/data/habitat_confirmations.xls) and summarized in Table \@ref(tab:tab-fish-site-090).

<br>

## Structure Remediation and Cost Estimate {-}

Structure replacement with an open bottomed culvert is recommended to provide access to the habitat located upstream of PSCIS crossing `r as.character(my_site)`.  Due to the large amount of fill on the road (`r my_pscis_info() %>% pull(fill_depth_meters)`) the size of a replacement bridge was estimated at `r my_pscis_info() %>% pull(recommended_diameter_or_span_meters)`m. The cost for work is estimated at \$`r as.character(my_cost_estimate() %>% pull(cost_est_1000s) * 1000)` for a cost benefit of \$`r as.character(my_cost_estimate() %>% pull(cost_net) * 1000)`/linear m and \$`r as.character(my_cost_estimate() %>% pull(cost_area_net) * 1000)`/m^2^.


## Conclusion {-}

There is an estimated `r my_priority_info() %>% pull(upstream_habitat_length_km)`km of mainstem habitat upstream of crossing `r as.character(my_site)` with habitat in the areas surveyed upstream of the crossing rated as high value for fry and juvenile salmonid rearing. However, fish sampling results indicate that westslope cutthrout trout are not currently utilizing the stream and densities of bull trout are very low.  This could be the result of the cold water conditions due to the positioning of the watershed at high elevation in the generally cold Rocky Mountain setting. Elk River FSR is a forest tenure liscensee road of the Ministry of Forests, Lands, Natural Resource Operations & Rural Development. The crossing was ranked as a `r my_priority_info() %>% pull(priority)` priority for proceeding to design for replacement with an open bottomed structure.

<br>

```{r tab-culvert-090, eval = T}
##had to pull this object out of the table push for some reason
print_tab_summary(site_photo_id = my_site_xref)

```

<br>

```{r tab-habitat-summary-090}
tab_hab_summary %>% 
  filter(Site  == my_site) %>% 
  select(-Site) %>% 
  my_kable(caption_text = paste0('Summary of habitat details for PSCIS crossing ', my_site, '.'))

```

<br>

```{r tab-fish-site-090, eval=T}
my_caption <- paste0('Electrofishing sites for PSCIS crossing ', my_site, '.')

tab_fish_site() %>% 
  my_kable(caption_text = my_caption)
```


<br>

```{r photo-090, fig.cap= my_caption, out.width = photo_width, eval=T}
my_photo = '20200917_123657_k_.JPG'

my_caption = paste0('Typical habitat downstream of PSCIS crossing ', my_site, '.')

grid::grid.raster(get_img(site = my_site_xref))

# knitr::include_graphics(get_img_path())
```

<br>

```{r photo-091, fig.cap= my_caption, out.width = photo_width, eval=T}

# 
my_photo = 'IMG_3058_k_.JPG'

my_caption = paste0('Typical habitat upstream of PSCIS crossing ', my_site, '.')

grid::grid.raster(get_img(site = my_site_xref))
# knitr::include_graphics(get_img_path())
```

<br>

```{r photo-092, fig.cap= my_caption, out.width = photo_width, eval=T}

# 
my_photo = '20200917_151420_k_d_f01_.JPG'

my_caption = paste0('Bull trout captured downstream of PSCIS crossing ', my_site, '.')

grid::grid.raster(get_img(site = my_site_xref))
# knitr::include_graphics(get_img_path())
```



