---
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup-50261, eval = F}
knitr::opts_chunk$set(echo=FALSE, message=FALSE, warning=FALSE)
source('R/packages.R')
source('R/functions.R')
source('R/functions-phase2.R')
source('R/tables-phase2.R')
source('R/tables.R')
source('R/extract-fish.R')
source('R/functions-fish.R')


```



```{r  load-data-50261}
my_site <- 50261

```


```{r eval = F}
my_bcfish <- my_bcfishpass() ##can be useful to look at

# `r paste0('# Appendix - Crossing ', my_site)`{-} 
# `r paste0('**', my_overview_info() %>% pull(road_name), ' - ', my_overview_info() %>% pull(stream_name), '**')`
```

`r appendix_title()`
`r appendix_subtitle()`



## Site Location {-}

Crossing `r as.character(my_site)` is located on a `r my_pscis_info() %>% pull(stream_name)`, approximately `r my_priority_info(loc = 'ds') %>% pull(survey_length_m)`m upstream from the confluence with the Elk River. The crossing is located on `r my_overview_info() %>% pull(road_name)` approximately 20km north of Sparwood. Elk Valley Park recreation site is located downstream of the crossing adjacent to the right bank of the river and the watershed upstream of the crossing is a habitat protection area with motor vehicle restrictions.


<br>

## Background {-}

At the crossing location, the stream is `r my_bcfishpass() %>% pull(stream_order) %>% english::ordinal()` order with a watershed area upstream of the road of approximately `r my_bcfishpass() %>% mutate(watershed_upstr_ha = watershed_upstr_ha/100) %>% pull(watershed_upstr_ha) %>% round(1)` km^2^ situated between Phillips Peak and Mount VanBuskirk. The elevation of the watershed ranges from a maximum of 2950 to 1250m at the culvert. Although, there are several modelled crossings on the mainstem of Hartley Creek upstream of the crossing, 2020 surveys indicated they were either fords or bridges. Modelled crossing 4606006 is a ford located approximately 1.2km upstream of the highway. There are several additional crossings modelled further upstream on the mainstem, however it is suspected that these crossings have been removed to facilitate the motor vehicle restrictions in place in the watershed. Westslope cutthroat trout have been recorded in `r my_pscis_info() %>% pull(stream_name)` downstream of PCSIS `r as.character(my_site)`.  Although there area multiple upstream survey sites recorded within provincial databases there have been no fish recorded upstream [@moeStreamInventorySample;@data_fish_obs]. 

<br>

PSCIS stream crossing `r as.character(my_site)` was ranked as a high priority for follow up with habitat confirmation due to the large amount of potential habitat modelled upstream (20km). The habitat confirmation was completed on `r my_pscis_info() %>% pull(date) %>% format(., format = "%B %d, %Y")`. A map of the watershed including areas surveyed is provided in Attachment 1 – Map [`r my_bcfishpass() %>% pull(map_tile_display_name)`](`r my_mapsheet()`).

<br>

## Stream Characteristics at Crossing {-}

At the time of the survey, the un-embedded and non-backwatered `r my_pscis_info() %>% pull(diameter_or_span_meters)`m diameter crossing was considered a barrier to upstream fish passage with a pipe length of `r my_pscis_info() %>% pull(length_or_width_meters)`m, a culvert slope of `r my_pscis_info() %>% pull(culvert_slope_percent)`%, a stream width ratio of `r my_pscis_info() %>% pull(stream_width_ratio)` and an outlet drop of `r my_pscis_info() %>% pull(outlet_drop_meters)`m (Table \@ref(tab:tab-culvert-50261)). Water temperature was `r my_habitat_info() %>% filter(rowname == 'temperature c') %>% pull(us)`$^\circ$C, pH was `r my_habitat_info() %>% filter(rowname == 'p h') %>% pull(us)` and conductivity was `r my_habitat_info() %>% filter(rowname == 'conductivity m s cm') %>% pull(us)`uS/cm.  

<br>


```{r eval=F}
##this is useful to get some comments for the report
hab_site %>% filter(site == my_site & location == 'ds') %>% pull(comments)
my_hab_info <- my_habitat_info()
```

## Stream Characteristics Downstream {-}

The stream was surveyed downstream from the culvert for `r my_priority_info(loc = 'ds') %>% pull(survey_length_m)`m. Overall total cover amount was rated as `r my_habitat_info2(loc = 'ds') %>% filter(rowname == 'total cover') %>% pull(v)` with  `r my_habitat_info2(loc = 'ds') %>% filter(v == 'dominant') %>% pull(rowname)` dominant. Cover was also present as `r my_habitat_info2(loc = 'ds') %>% filter(v == 'sub-dominant') %>% pull(rowname) %>% knitr::combine_words()` (Table \@ref(tab:tab-habitat-summary-50261), Figure \@ref(fig:photo-50261-01)). The average channel width was `r my_habitat_info3(loc = 'ds', row = 'avg channel width m')`m, the average wetted width was `r my_habitat_info3(loc = 'ds', row = 'avg wetted width m')`m and the average gradient was `r my_habitat_info3(loc = 'ds', row = 'average gradient percent')`%.  The dominant substrate was `r my_habitat_info3(loc = 'ds', row = "bed material dominant")` with `r my_habitat_info3(loc = 'ds', row = "bed material subdominant")` subdominant.  Pools were present to 0.8m deep and gravels/cobbles suitable for resident and fluvial spawning were present.  A ford crosses the stream at the Fortis gasline located approximately 115m downstream of the highway. Overall, habitat was rated as `r my_priority_info(loc = 'ds') %>%  pull(hab_value)` value providing an important migration corridor and habitat suitable for resident and fluvial salmonid rearing and spawning.

<br>



```{r eval=F}
##this is useful to get some comments for the report
hab_site %>% filter(site == my_site & location == 'us') %>% pull(comments)
my_hab_info <- my_habitat_info()
```


<br>

## Stream Characteristics Upstream {-}

The stream was surveyed immediately upstream from the culvert for `r my_priority_info(loc = 'ds') %>% pull(survey_length_m)`m. Overall, total cover amount was rated as `r my_habitat_info2(loc = 'us') %>% filter(rowname == 'total cover') %>% pull(v)` with  `r my_habitat_info2(loc = 'us') %>% filter(v == 'dominant') %>% pull(rowname)` dominant. Cover was also present as `r my_habitat_info2(loc = 'us') %>% filter(v == 'sub-dominant') %>% pull(rowname) %>% knitr::combine_words()` (Table \@ref(tab:tab-habitat-summary-50261), Figure \@ref(fig:photo-50261-02)). The average channel width was `r my_habitat_info3(loc = 'us', row = 'avg channel width m')`m, the average wetted width was `r my_habitat_info3(loc = 'us', row = 'avg wetted width m')`m and the average gradient was `r my_habitat_info3(loc = 'us', row = 'average gradient percent')`%.  The dominant substrate was `r my_habitat_info3(loc = 'us', row = "bed material dominant")` with `r my_habitat_info3(loc = 'us', row = "bed material subdominant")` subdominant. Habitat became increasingly complex with increased distance upstream.  Throughout the area surveyed there were occasional pools available with depths suitable for juvenile/adult salmonid overwintering and rearing and frequent sections of gravels available for resident and fluvial salmonid spawning. Overall, habitat was rated as `r my_priority_info(loc = 'ds') %>%  pull(hab_value)` value.  

<br>


## Structure Remediation and Cost Estimate {-}

Structure replacement with an open bottomed structure is recommended to provide access to the habitat located upstream of PSCIS crossing `r as.character(my_site)`.  The cost of the work is estimated at \$`r as.character(my_cost_estimate() %>% pull(cost_est_1000s) * 1000)` for a cost benefit of \$`r as.character(my_cost_estimate() %>% pull(cost_net) * 1000)`/linear m and \$`r as.character(my_cost_estimate() %>% pull(cost_area_net) * 1000)`/m^2^.

<br>

## Conclusion {-}

There is an estimated `r my_priority_info() %>% pull(upstream_habitat_length_km)`km of mainstem habitat available upstream of crossing `r as.character(my_site)` with habitat rated as `r my_priority_info(loc = 'us') %>% pull(hab_value)` value, representing an important migration corridor with high rearing and spawning potential for resident and fluvial salmonids. `r my_overview_info() %>% pull(road_name)` is the responsiblity of the Ministry of Transportation and Infrastructure. The crossing was ranked as a `r my_priority_info() %>% pull(priority)` priority for proceeding to design for replacement with an open bottomed structure. 

\pagebreak

```{r tab-culvert-50261, eval = T}
print_tab_summary()

```

\pagebreak

```{r tab-habitat-summary-50261}
tab_hab_summary %>% 
  filter(Site  == my_site) %>% 
  select(-Site) %>%
  my_kable(caption_text = paste0('Summary of habitat details for PSCIS crossing ', my_site, '.'))

```



<br>

```{r photo-50261-01, fig.cap= my_caption, out.width = photo_width, eval=T}
my_photo = pull_photo_by_str(str_to_pull = '_d1')

my_caption = paste0('Habitat downstream of PSCIS crossing ', my_site, '.')

grid::grid.raster(get_img())

# knitr::include_graphics(get_img_path())
```

<br>

```{r photo-50261-02, fig.cap= my_caption, out.width = photo_width, eval=T}
my_photo = pull_photo_by_str(str_to_pull = '_u1')

my_caption = paste0('Habitat upstream of PSCIS crossing ', my_site, '.')

grid::grid.raster(get_img())

# knitr::include_graphics(get_img_path())
```


