
```{r setup-04, eval = F}
knitr::opts_chunk$set(echo=FALSE, message=FALSE, warning=FALSE)
source('R/packages.R')
source('R/functions.R')
source('R/functions-phase2.R')
source('R/tables-phase2.R')
source('R/tables.R')
source('R/extract-fish.R')
source('R/functions-fish.R')


```



```{r  load-data-04}
my_site <- 62516
```


`r paste0('# Appendix - Crossing ', my_site)`{-} 
`r paste0('**', my_overview_info() %>% pull(road_name), ' - ', my_overview_info() %>% pull(stream_name), '**')`


## Site Location {-}

Crossing `r as.character(my_site)` is located on a `r my_pscis_info() %>% pull(stream_name)`, approximately `r my_priority_info(loc = 'ds') %>% pull(survey_length_m)`m upstream from the confluence with Lizard Creek. This culvert has also been recorded in PSCIS as crossing 50153. The culvert is located at 2.5km on Island Lake Lodge Road which is an extension of Mt.Fernie Park Road accessed from Highway 3 within Fernie city limits.  The Mt.Fernie Provincial Park is a popular recreational destination for hikers, sightseers and mountain bikers.  Island Lake Lodge is located at 1400m of elevation near Island Lake and is a year round tourist destination providing accommodations, guided hiking and backcountry catskiing. The stream has been diverted from its historic channel and runs adjacent to a historic road to Lizard Creek approximately 500 downstream of the location of the confluence in the freshwater atlas.  


<br>

## Background {-}

At the crossing location, the stream had good flow and is located within an area of old growth cedar adjacent to a recreation trail. At the time of the survey the stream was the highest volume tributary to Lizard Creek located on the east side of the valley. No fisheries information was available for the stream [@moeStreamInventorySample]. Downstream however, Lizard Creek supports westslope cutthroat trout, bull trout, mountain whitefish, brook trout, longnose sucker and longnose dace [@data_fish_obs]. @elkriveralliance2020ElkRiver conducted redd surveys in Lizard Creek in 2019 with a total of 55 redds observed within a 2.4km of stream surveyed comprising the largest densities of redds of the four tributaries surveyed (22.9 redds/km).

<br>

PSCIS stream crossing `r as.character(my_site)` was ranked as a high priority for follow up with habitat confirmation due to the relatively large size of the stream recorded in PSCIS (channel width = 3.5m) relative to other tributary streams in the Lizard Creek watershed and because it was rated as containing high value habitat by @vastFishPassage2013 and @grainger2011FishPassage2011. There is a bridge (PSCIS 197566) located on the Lazy Lizard recreation trail approximatley 350m upstream of the crossing and an unassessed modelled crossing (modelled ID 4600929) approximately 950m upstream.  Downstream approximately 400m there is a foot bridge (PSCIS 197544) on a recreational hiking/biking trail.  The habitat confirmation was completed on `r my_pscis_info() %>% pull(date) %>% format(., format = "%B %d, %Y")`. A map of the watershed including areas surveyed is provided in Attachment 1 – Map [`r my_bcfishpass() %>% pull(map_tile_display_name)`](`r my_mapsheet()`).

<br>

## Stream Characteristics at Crossing {-}

Th culvert located at PSCIS `r as.character(my_site)` appeared to have been replaced in 2020 with fresh rock and road fill present.  At the time of the survey, the un-embedded and non-backwatered `r my_pscis_info() %>% pull(diameter_or_span_meters)`m diameter crossing was considered a barrier to upstream fish passage with a pipe length of `r my_pscis_info() %>% pull(length_or_width_meters)`m, a culvert slope of `r my_pscis_info() %>% pull(culvert_slope_percent)`%, a stream width ratio of `r my_pscis_info() %>% pull(stream_width_ratio)` and an outlet drop of `r my_pscis_info() %>% pull(outlet_drop_meters)`m (Table \@ref(tab:tab-culvert-040)). Water temperature was `r my_habitat_info() %>% filter(rowname == 'temperature c') %>% pull(us)`$^\circ$C, pH was `r my_habitat_info() %>% filter(rowname == 'p h') %>% pull(us)` and conductivity was `r my_habitat_info() %>% filter(rowname == 'conductivity m s cm') %>% pull(us)`uS/cm.  

<br>


```{r eval=F}
##this is useful to get some comments for the report
hab_site %>% filter(site == my_site & location == 'us') %>% pull(comments)
my_priority <-  my_priority_info()

```

## Stream Characteristics Downstream {-}

The stream was surveyed downstream from the culvert for `r my_priority_info(loc = 'ds') %>% pull(survey_length_m)`m to Lizard Creek. Overall, total cover amount was rated as `r my_habitat_info2(loc = 'ds') %>% filter(rowname == 'total cover') %>% pull(v)` with  `r my_habitat_info2(loc = 'ds') %>% filter(v == 'dominant') %>% pull(rowname)` dominant. Cover was also present as `r my_habitat_info2(loc = 'ds') %>% filter(v == 'sub-dominant') %>% pull(rowname) %>% knitr::combine_words()` (Table \@ref(tab:tab-habitat-summary-040), Figure \@ref(fig:photo-040)). The average channel width was `r my_habitat_info3(loc = 'ds', row = 'avg channel width m')`m, the average wetted width was `r my_habitat_info3(loc = 'ds', row = 'avg wetted width m')`m and the average gradient was `r my_habitat_info3(loc = 'ds', row = 'average gradient percent')`%.  The dominant substrate was `r my_habitat_info3(loc = 'ds', row = "bed material dominant")` with `r my_habitat_info3(loc = 'ds', row = "bed material subdominant")` subdominant. Adjacent to the historic road, on the right bank of stream, there was very limited shrub and tree riparian vegetation. Downstream of the crossing there were frequent sections of gravels suitable for salmonid spawning and no barriers or obstacles to fish passage. Habitat was rated as `r my_priority_info(loc = 'ds') %>%  pull(hab_value)` value for fry/juvenile salmonid rearing.  

<br>

## Stream Characteristics Upstream {-}

The stream was surveyed upstream from the culvert for `r my_priority_info(loc = 'us') %>% pull(survey_length_m)`m. Overall, total cover amount was rated as `r my_habitat_info2(loc = 'us') %>% filter(rowname == 'total cover') %>% pull(v)` with  `r my_habitat_info2(loc = 'us') %>% filter(v == 'dominant') %>% pull(rowname)` dominant. Cover was also present as `r my_habitat_info2(loc = 'us') %>% filter(v == 'sub-dominant') %>% pull(rowname) %>% knitr::combine_words()` (Table \@ref(tab:tab-habitat-summary-040), Figure \@ref(fig:photo-040)). The average channel width was `r my_habitat_info3(loc = 'us', row = 'avg channel width m')`m, the average wetted width was `r my_habitat_info3(loc = 'us', row = 'avg wetted width m')`m and the average gradient was `r my_habitat_info3(loc = 'us', row = 'average gradient percent')`%.  There were abundant gravels suitable for resident westslope cutthrout trout spawning throughout (Figure \@ref(fig:photo-040)).  Fry were observed within the area surveyed to 540m upstream of the culvert where gradients increased to >20% for a distance of approximately 15m.  Although no sampling was conducted, no fish were observed above this high gradient section. Habitat value was rated as `r my_priority_info(loc = 'us') %>% pull(hab_value)` for fry/juvenile salmonid rearing. 


<br>

## Structure Remediation and Cost Estimate {-}

Structure replacement with an open bottomed structure is recommended to provide access to the habitat located upstream of PSCIS crossing `r as.character(my_site)`.  The cost of the work is estimated at \$`r as.character(my_cost_estimate() %>% pull(cost_est_1000s) * 1000)` for a cost benefit of \$`r as.character(my_cost_estimate() %>% pull(cost_net) * 1000)`/linear m and \$`r as.character(my_cost_estimate() %>% pull(cost_area_net) * 1000)`/m^2^.

<br>


## Conclusion {-}

There is `r my_priority_info() %>% pull(upstream_habitat_length_km)`km of mainstem habitat upstream of crossing `r as.character(my_site)` with habitat rated as high value for fry and juvenile salmonid rearing. The road may be part of the Island Lake Recreational tenure or solely the responsiblity of the Ministry of Forests, Lands, Natural Resource Operations & Rural Development. The crossing was ranked as a `r my_priority_info() %>% pull(priority)` priority for proceeding to design for replacement.

<br>

```{r tab-culvert-040, eval = T}
print_tab_summary()
```

<br>

```{r tab-habitat-summary-040}
tab_hab_summary %>% 
  filter(Site  == my_site) %>% 
  select(-Site) %>% 
  my_kable(caption_text = paste0('Summary of habitat details for PSCIS crossing ', my_site, '.'))

```


<br>

```{r photo-040, fig.cap= my_caption, out.width = photo_width, eval=T}
my_photo = 'IMG_3366_k_.JPG'

my_caption = paste0('Typical habitat downstream of PSCIS crossing ', my_site, '.')

grid::grid.raster(get_img())

# knitr::include_graphics(get_img_path())
```

<br>

```{r photo-041, fig.cap= my_caption, out.width = photo_width,eval=T}
my_photo = '20200923_115923_k_.JPG'

my_caption = paste0('Typical habitat upstream of PSCIS crossing ', my_site, '.')

grid::grid.raster(get_img())
# knitr::include_graphics(get_img_path())
```

<br>

```{r photo-042, fig.cap= my_caption, out.width = photo_width, eval=T}
my_photo = '20200923_115933_u_g_.JPG'

my_caption = paste0('Gravels suitable for spawning upstream of PSCIS crossing ', my_site, '.')

grid::grid.raster(get_img())
# knitr::include_graphics(get_img_path())
```

