--- 
title: "Elk River Watershed Group Fish Passage Restoration Planning"
author:
 - Prepared for Canadian Wildlife Federation
 - Prepared by Al Irvine, B.Sc., R.P.Bio.
 - New Graph Environment
 - Version 0.0.3
date: "2020-12-31 DRafT"
toc-title: Table of Contents
site: bookdown::bookdown_site
output: bookdown::gitbook
documentclass: book
bibliography: [book.bib, packages.bib]
biblio-style: apalike
link-citations: no
github-repo: rstudio/bookdown-demo
description: "My Description."


---


```{r setup, include = TRUE, echo =FALSE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo=FALSE, message=FALSE, warning=FALSE, dpi=60, out.width = "100%")

##define a picture specific output width you can put in chunk headers
photo_width <- "100%"
# photo_width <- "80%"
# ##we need a switch to turn off html only so the printed version looks ok
html_on <- TRUE
# html_on <- FALSE
# 


##try to fix the kable caption size https://stackoverflow.com/questions/45018397/changing-the-font-size-of-figure-captions-in-rmarkdown-html-output

# fig.align="center", out.width = "90%",
source('R/packages.R')
source('R/functions.R')
source('R/functions-phase2.R')
source('R/tables.R')
source('R/tables-phase2.R')
source('R/tables-phase1-cost-estimate.R')
source('R/tables-phase2-cost-estimate.R') ##the order of these matters to much - should fix.  need to make cost estimate table seperate
source('R/tables-reporting.R')
source('R/extract-fish.R')
source('R/functions-fish.R')

##set the font default size for the tables.  had a crash and think this might have been it so put after functions.R
##seems odd but not risking again for now.
font_set <- 11
# font_set <- 9


```

```{r include=FALSE}
# automatically create a bib database for R packages
knitr::write_bib(c(
  .packages(), 'bookdown', 'knitr', 'rmarkdown'
), 'packages.bib')
```


# Executive Summary {.front-matter .unnumbered}

New Graph Environment and Nupqu Limited Partnership were retained by the Canadian Wildlife Federation in the fall of 2020 to plan and conduct fish passage and habitat confirmation assessments at road-stream crossings within the Elk River watershed group.  Although some planning for future assessments was conducted for both the Elk River watershed upstream of the Elko Dam near Elko, BC and the Flathead River, on the ground surveys focused on the Elk River and tributaries located upstream of the Elko Dam. 

<br>

A total of `r n_distinct(pscis$my_crossing_reference)` phase 1 assessments were conducted with `r pscis %>% filter(barrier_result == 'Passable') %>% nrow()` crossings considered "passable", `r pscis %>% filter(barrier_result == 'Potential') %>% nrow()` crossings considered 'potential' barriers and `r pscis %>% filter(barrier_result == 'Barrier') %>% nrow()` crossings considered barriers.

<br>

Habitat confirmation assessments were conducted at `r hab_site_priorities %>% nrow()` sites with a total of approximately 15 km of stream assessed. `r hab_site_priorities %>% filter(priority %ilike% 'high') %>% nrow() %>% english::as.english() %>% str_to_title()` crossings were rated as high priorities for proceeding to design for replacement, `r hab_site_priorities %>% filter(priority %ilike% 'Moderate') %>% nrow() %>% english::as.english() %>% str_to_title()` crossings were rated as moderate priorities for proceeding to design for replacement, `r hab_site_priorities %>% filter(priority %ilike% 'Low') %>% nrow() %>% english::as.english()` crossings were rated as a low priority and `r hab_site_priorities %>% filter(priority %ilike% 'no fix') %>% nrow() %>% english::as.english()` rated as "no fix".

