--- 
title: "Bulkley River and Morice River Watershed Groups Fish Passage Restoration Planning"
author:
 - Prepared for Habitat Conservation Trust Fund, BC Fish Passage Remediation Program and the Canadian Wildlife Federation
 - Prepared by Al Irvine, B.Sc., R.P.Bio.
 - New Graph Environment on behalf of Society for Ecosytem Restoration Northern BC
 - Version DRafT - 0.0.1
date: |
   `r format(Sys.Date(), "%Y-%m-%d")`
toc-title: Table of Contents
site: bookdown::bookdown_site
output: bookdown::gitbook
documentclass: book
bibliography: [book.bib, packages.bib]
biblio-style: apalike
link-citations: no
github-repo: rstudio/bookdown-demo
description: "My Description."

---


```{r setup, include = TRUE, echo =FALSE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo=FALSE, message=FALSE, warning=FALSE, dpi=60, out.width = "100%")

##define a picture specific output width you can put in chunk headers
##we need a switch to turn off html only so the printed version looks ok
photo_width <- "100%"
html_on <- TRUE

##these are the settings for the pagedown version!! DONT FORGET THE FONT DOWN BELOW TOO!!!!!!!!!!!!!!
# photo_width <- "80%"
# html_on <- FALSE



##try to fix the kable caption size https://stackoverflow.com/questions/45018397/changing-the-font-size-of-figure-captions-in-rmarkdown-html-output

# fig.align="center", out.width = "90%",
source('R/packages.R')
source('R/functions.R')
source('R/functions-phase2.R')
source('R/0310-tables.R')
source('R/0320-tables-phase2.R')
source('R/0330-tables-phase1-cost-estimate.R')
source('R/0340-tables-phase2-cost-estimate.R') ##the order of these matters to much - should fix.  need to make cost estimate table seperate
source('R/0350-tables-reporting.R')
source('R/0380-extract-fish-to-hab-con-density-summary.R')
source('R/functions-fish.R')


# ##made this with 0380-extract-fish-to-hab-con.R
# hab_fish_dens <- readr::read_csv(file = paste0(getwd(), '/data/extracted_inputs/hab_fish_dens.csv'))

##set the font default size for the tables.  had a crash and think this might have been it so put after functions.R
##seems odd but not risking again for now.
font_set <- 11
# font_set <- 9


```

```{r include=FALSE}
# automatically create a bib database for R packages
knitr::write_bib(c(
  .packages(), 'bookdown', 'knitr', 'rmarkdown'
), 'packages.bib')
```


# Executive Summary {.front-matter .unnumbered}

New Graph Environment was retained by the Habitat Conservation Trust, the Provincial Fish Passage Remediation Program and the Canadian Wildlife Federation in 2020 to plan and conduct fish passage and habitat confirmation assessments at road-stream crossings as part of connectivity restoration planning. Although planning was conducted for both the Morice River watershed group and the Bulkley River watershed group, on the ground surveys in 2020 focused primarily within Bulkley River tributaries. 

<br>

A total of `r n_distinct(pscis_phase1_reassessments$my_crossing_reference)` Phase 1 assessments were conducted with `r pscis_phase1_reassessments %>% filter(barrier_result == 'Passable') %>% nrow()` crossings considered "passable", `r pscis_phase1_reassessments %>% filter(barrier_result == 'Potential') %>% nrow() %>% english::as.english()` crossings considered "potential" barriers and `r pscis_phase1_reassessments %>% filter(barrier_result == 'Barrier') %>% nrow()` crossings considered "barriers" according to threshold values based on culvert embedment, outlet drop, slope, diameter (relative to channel size) and length.

<br>

Habitat confirmation assessments were conducted at `r hab_site_priorities %>% nrow()` sites with a total of approximately 18 km of stream assessed. `r hab_site_priorities %>% filter(priority %ilike% 'high') %>% nrow() %>% english::as.english() %>% str_to_title()` crossings were rated as high priorities for proceeding to design for replacement, `r hab_site_priorities %>% filter(priority %ilike% 'Moderate') %>% nrow() %>% english::as.english()` crossings were rated as moderate priorities for proceeding to design for replacement, `r hab_site_priorities %>% filter(priority %ilike% 'Low') %>% nrow() %>% english::as.english()` crossings were rated as a low priority.

